### 2016-03-16 paper/paper.tex
initial branch commit contains the paper and references - including the article.tex from blackboard (article_bb) - including the article.tex from the previous year from blackboard, which might contains changes (article_old_tim) -including the paper.tex from my (Tim) previous project (paper_old_tim) -including the references from paper_old_tim (references_old_tim)

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[twoside]{article}

\usepackage{amsmath,amssymb,amsthm} % Mathematical Symbols, styles, etc

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}

\usepackage[sc]{mathpazo} % Use the Palatino font
\usepackage[T1]{fontenc} % Use 8-bit encoding that has 256 glyphs
\linespread{1.05} % Line spacing - Palatino needs more space between lines
\usepackage{microtype} % Slightly tweak font spacing for aesthetics

\usepackage[hmarginratio=1:1,top=32mm,columnsep=20pt]{geometry} % Document margins
\usepackage{multicol} % Used for the two-column layout of the document
\usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption} % Custom captions under/above floats in tables or figures
\usepackage{booktabs} % Horizontal rules in tables
\usepackage{float} % Required for tables and figures in the multi-column environment - they need to be placed in specific locations with the [H] (e.g. \begin{table}[H])
\usepackage{hyperref} % For hyperlinks in the PDF

\usepackage{lettrine} % The lettrine is the first enlarged letter at the beginning of the text
\usepackage{paralist} % Used for the compactitem environment which makes bullet points with less space between them

\usepackage{abstract} % Allows abstract customization
\renewcommand{\abstractnamefont}{\normalfont\bfseries} % Set the "Abstract" text to bold
\renewcommand{\abstracttextfont}{\normalfont\small\itshape} % Set the abstract itself to small italic text

\usepackage{titlesec} % Allows customization of titles
%\renewcommand\thesection{\Roman{section}} % Roman numerals for the sections
%\renewcommand\thesubsection{\Roman{subsection}} % Roman numerals for subsections
\titleformat{\section}[block]{\large\scshape}{\thesection.}{1em}{} % Change the look of the section titles
\titleformat{\subsection}[block]{\large}{\thesubsection.}{1em}{} % Change the look of the section titles

\usepackage{fancyhdr} % Headers and footers
\pagestyle{fancy} % All pages have headers and footers
\fancyhead{} % Blank out the default header
\fancyfoot{} % Blank out the default footer
\fancyhead[C]{W.G. Puttenstein, R.L.H. Fontein, H.D. van Wieren, T. Kerkhoven: \shorttitle} % Custom header text
\fancyfoot[RO,LE]{\thepage} % Custom footer text

% Bibliography
\usepackage[backend=bibtex, sorting=none]{biblatex}
\bibliography{references.bib}

%----------------------------------------------------------------------------------------
%	TITLE SECTION
%----------------------------------------------------------------------------------------

\newcommand{\articletitle}{Comparison between Time Complexity of Different Color Refinement Algorithms in Graph Isomorphism}
\newcommand{\shorttitle}{Color Refinement Time Complexity in Graph Isomorphism}

\title{\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{\articletitle}} % Article title

\author{
\large
\textsc{Gerwin Puttenstein, Rick Fontein, Huub van Wieren, and Tim Kerkhoven}\\[2mm] % Your name
\normalsize University of Twente \\ % Your institution
\normalsize \href{mailto:w.g.puttenstein@student.utwente.nl}{w.g.puttenstein@student.utwente.nl}, 
\href{mailto:r.l.h.fontein@student.utwente.nl}{r.l.h.fontein@student.utwente.nl},\\
\normalsize \href{mailto:h.d.vanwieren@student.utwente.nl}{h.d.vanwieren@student.utwente.nl},
\href{mailto:t.kerkhoven@student.utwente.nl}{t.kerkhoven@student.utwente.nl}% Your email addresses
}

\date{\today}

%----------------------------------------------------------------------------------------

\begin{document}

\thispagestyle{empty}
\maketitle % Insert title

%----------------------------------------------------------------------------------------
%	ABSTRACT
%----------------------------------------------------------------------------------------

\begin{abstract}

\noindent Test

\end{abstract}

%----------------------------------------------------------------------------------------
%	ARTICLE CONTENTS
%----------------------------------------------------------------------------------------

\begin{multicols}{2} % Two-column layout throughout the main article text

%------------------------------------------------
\section{Introduction}
This paper will compare the time complexity of two color refinement algorithms and confirm them with computational results. The first algorithm uses the color refinement algorithm as shown by J. de Jong\cite{presentation:slidesPartI}, while the other algorithm is based on DFA minimization as shown by P. Bonsma\cite{presentation:slidesPartIII}. The algorithms have been implemented and their computation times on test sets will be used to confirm the time complexity.

Color refinement algorithms are used in solving the graph isomorphism problem, which is the computational problem of determining whether two finite graphs are isomorphic. It is one of a very small number of problems belonging to NP neither known to be solvable in polynomial time nor NP-complete: it is one of 12 such problems listed by Garey \& Johnson (1979)\cite{book:gareyJohnson1979}, and the only one of that list whose complexity remains unresolved.\cite{website:wikiGI}

%------------------------------------------------
\section{Scope \& Objectives}

%------------------------------------------------
\section{Methods}

%------------------------------------------------
\section{Algorithms} %Descriptions + time complexity

%------------------------------------------------
\section{Results}

%------------------------------------------------
\section{Conclusions}

%----------------------------------------------------------------------------------------
%	REFERENCE LIST
%----------------------------------------------------------------------------------------

\printbibliography
%bibliography{references}

%----------------------------------------------------------------------------------------
%	APPENDICES
%----------------------------------------------------------------------------------------

%----------------------------------------------------------------------------------------

\end{multicols}
%\begin{appendices}

%----------------------------------------------------------------------------------------

%\end{appendices}
\end{document}

### 2016-03-16 paper/paper_old_tim.tex
initial branch commit contains the paper and references - including the article.tex from blackboard (article_bb) - including the article.tex from the previous year from blackboard, which might contains changes (article_old_tim) -including the paper.tex from my (Tim) previous project (paper_old_tim) -including the references from paper_old_tim (references_old_tim)

 \documentclass[twoside]{article}

\usepackage{amsmath,amssymb,amsthm} % Mathematical Symbols, styles, etc

\usepackage[utf8]{inputenc} % UTF-8 character encoding stuff
\usepackage[toc,page]{appendix} % appendix

\usepackage{array}
\newcolumntype{C}[1]{>{\centering\arraybackslash}p{#1}}

% Used to enable use of the defined labeled enunciations. Usage: 
%\begin{definition}text\end{definition}
\theoremstyle{definition}
\newtheorem{definition}{Definition}

\theoremstyle{plain}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}

% Python code
\usepackage{listings}
\lstset{
numbers=left,
breaklines=true,
tabsize=4
}

% --Optional-- Font and tweaking used in template
\usepackage[sc]{mathpazo} % Use the Palatino font
\usepackage[T1]{fontenc} % Use 8-bit encoding that has 256 glyphs
\linespread{1.05} % Line spacing - Palatino needs more space between lines
\usepackage{microtype} % Slightly tweak font spacing for aesthetics

% Formatting, refs and stuff
\usepackage[hmarginratio=1:1,top=32mm,columnsep=20pt]{geometry} % Document margins
\usepackage{multicol} % Used for the two-column layout of the document
\usepackage[hang, small,labelfont=bf,up,textfont=it,up]{caption} % Custom captions under/above floats in tables or figures
\usepackage{booktabs} % Horizontal rules in tables
\usepackage{float} % Required for tables and figures in the multi-column environment - they need to be placed in specific locations with the [H] (e.g. \begin{table}[H])
\usepackage{hyperref} % For hyperlinks in the PDF

% Fancy lists. Usage: http://lmgtfy.com/?q=latex+paralist
\usepackage{paralist} % Used for the compactitem environment which makes bullet points with less space between them

% Abstract customization
\usepackage{abstract} % Allows abstract customization
\renewcommand{\abstractnamefont}{\normalfont\bfseries} % Set the "Abstract" text to bold
\renewcommand{\abstracttextfont}{\normalfont\small\itshape} % Set the abstract itself to small italic text

% Title customization
\usepackage{titlesec} % Allows customization of titles
%\renewcommand\thesection{\Roman{section}} % Roman numerals for the sections
%\renewcommand\thesubsection{\Roman{subsection}} % Roman numerals for subsections
\titleformat{\section}[block]{\large\scshape}{\thesection.}{1em}{} % Change the look of the section titles
\titleformat{\subsection}[block]{\large}{\thesubsection.}{1em}{} % Change the look of the section titles

% Header and footer customization
\usepackage{fancyhdr} % Headers and footers
\pagestyle{fancy} % All pages have headers and footers
\fancyhead{} % Blank out the default header
\fancyfoot{} % Blank out the default footer
\fancyhead[C]{T.\ Kemp, T.\ Kerkhoven, J.\ Klein Brinke, T.\ Sonderen: \shorttitle} % Custom header text
\fancyfoot[RO,LE]{\thepage} % Custom footer text

% Bibliography
\usepackage[backend=bibtex, sorting=none]{biblatex}
\bibliography{references_old_tim.bib}

%----------------------------------------------------------------------------------------
%	TITLE SECTION
%----------------------------------------------------------------------------------------

\newcommand{\articletitle}{The Effects of Preprocessing Methods on Computation Times for Graph Isomorphisms}
\newcommand{\shorttitle}{Preprocessing for Graph Isomorphisms}

\title{\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{\articletitle}} % Article title

\author{
\large
\textsc{Tim Kemp, Tim Kerkhoven, Jeroen Klein Brinke, and Tim Sonderen}\\[2mm] % Your names
\normalsize University of Twente \\ % Your institution
\normalsize \href{mailto:t.kerkhoven@student.utwente.nl}{t.kerkhoven@student.utwente.nl}, 
\href{mailto:t.kemp@student.utwente.nl}{t.kemp@student.utwente.nl},\\ \normalsize
\href{mailto:j.klein.brinke@student.utwente.nl}{j.klein.brinke@student.utwente.nl}, \href{mailto:t.sonderen@student.utwente.nl}{t.sonderen@student.utwente.nl}% Your email addresses
}

\date{\today}

%----------------------------------------------------------------------------------------

\begin{document}

\thispagestyle{empty}
\maketitle % Insert title

%----------------------------------------------------------------------------------------
%	ABSTRACT
%----------------------------------------------------------------------------------------

\begin{abstract}

This research paper shows how a certain preprocessing method, namely the finding and replacing of (false) twins, can be used to decrease the computation time on the graph isomorphism (GI) problem. The hypothesis of this study was that finding and replacing sets of (false) twins by a single colored node will decrease the computation time on finding isomorphisms between graphs. The paper also contains a mathematical proof which supports the claim that (false) twins can be replaced by a single colored node. During the project, an algorithm has been developed to find and replace these (false) twins.  The algorithm computed different types of graphs with preprocessing both on and off. The results of these computational experiments have been compared and based on the differences in computation time a conclusion has been drawn. This conclusion is that it heavily depends on the type of graph: the more (false) twins a graph has, the faster a graph can be processed. The time of finding all the (false) twins can be neglected. The results also showed that for some graphs for which the GI problem cannot be solved in reasonable time without this preprocessing method, they could be solved with this preprocessing method.

\end{abstract}

%----------------------------------------------------------------------------------------
%	ARTICLE CONTENTS
%----------------------------------------------------------------------------------------

\begin{multicols}{2} % Two-column layout throughout the main article text

%------------------------------------------------
\section{Introduction}

This paper will discuss the effect of a certain preprocessing method on the computation time of the graph isomorphism problem. This paper will mostly focus on the algorithms and their correctness, in combination with computational experiments and their results.

The graph isomorphism problem is the computational problem of determining whether two finite graphs are isomorphic. Besides its practical importance, the graph isomorphism problem is a curiosity in computational complexity theory as it is one of the few problems belonging to NP neither known to be solvable in polynomial time nor NP-complete: it is one of only 12 such problems listed by Garey \& Johnson (1979)\cite{book:garyJohnson1979}, and one of only two on that list whose complexity remains unresolved. As of 2008 the best algorithm (Eugene Luks, 1983) has run time $2^{O(\sqrt{n log n})}$ for graphs with \emph{n} vertices.\cite{article:davidJohnson2005}\cite{inproceedings:babaiCodenotti2008}\cite{website:wikiGI}

The preprocessing part consists of algorithms that look at the graphs before determining whether they are isomorphic with the intent of finding some characteristic that might be used to speed up the actual process of the determining whether the graphs are isomorphic.

%------------------------------------------------

\section{Scope \& Objectives}

The aim of this research paper is to determine whether a certain preprocessing method significantly reduces the time required to check whether two graphs are isomorphic. At the core, two cases will be compared: a basic case without preprocessing and a case with preprocessing.

For this paper, the time complexity of the algorithms will be calculated. The difference in performance between the cases will be shown by computational results.

This paper will give a description of the algorithms used for preprocessing, as well as proof of correctness. The algorithms used to check for isomorphisms will be left out of consideration.

%------------------------------------------------
\section{Methods}
To determine whether the algorithms will speed up the process of finding graph isomorphisms, a computational experiment was performed. In this experiment, the algorithms were used on a variety of differently shaped graphs, and sets of graphs, in conjunction with an algorithm which checks for isomorphisms.

\subsection{Experiment Setup}
For the testing, a single computer was used to collect all the timed data. The tests were run in JetBrains' PyCharm, using the built-in functions to run the python code. The python interpreter used was Python 3.4 and the computer was running Windows 10 Pro Technical Preview (at 3.75 GHz (average speed, max 3.9 GHz) and 8.00 GB) at the time of the testing. Efforts were made to ensure all the testing was performed under the same circumstances.

\subsection{Testing}
In the experiment, two cases were tested: a basic case, without preprocessing, and a case with preprocessing. In both cases a timer was added to the code. The timer started before any other piece of code and ended after the algorithms were completely finished. This way all the required computations are timed, including the loading of the graph, the additional time used to run preprocessing algorithms and the finding of isomorphisms.

For every chosen test instance, the run time of both algorithms was determined. After all instances were computed, the results were compared.

%------------------------------------------------
\section{Description of Algorithms}
For hard instances of the graph isomorphism problem, the coarsest stable coloring of a graph still has colors which are not unique. Two structures for which this is the case are twins and false twins, they are defined in definition \ref{deftwins} and \ref{deffalsetwins}.
\begin{definition}
Let $G = (V,E)$ be a graph. The \emph{neighbourhood} $N(v)$ of $v \in V$ is:
$N(v) = \left\{{u \in V : \{u,v\} \in E}\right\}$
\end{definition}
\begin{definition}\label{deftwins}
Let $G = (V,E)$ be a graph.
Two vertices $u, v \in V$ are \emph{twins} if $\{u,v\} \in E$ and $N(u)\textbackslash \{v\} = N(v)\textbackslash \{u\}$.
\end{definition}
\begin{definition}\label{deffalsetwins}
Let $G = (V,E)$ be a graph.
Two vertices $u, v \in V(G)$ are \emph{false twins} if $N(u) = N(v).$
\end{definition}

To solve the graph isomorphism problem more efficiently, we use preprocessing algorithms to replace twins and false twins by single precolored vertices such that:
\begin{compactenum}
\item Vertices in the new graph have the same color $\iff$ The corresponding twinsets or false twinsets in the old graph have the same size (number of vertices).
\item Two vertices in the new graph cannot have the same color if one of them corresponds to a twinset and the other one two a false twinset in the old graph.
\end{compactenum}

\subsection{Implementation}
The algorithm is divided into two subroutines. The subroutine \emph{gettwins} finds all twins and false twins and the subroutine \emph{preprocessing} replaces each (false) twin by a single colored vertex as described above. In the following subsections these algorithms and their worst case time complexities will be analyzed. The elementary operation is the comparison of two elements. In those subsections $n$ is the number of vertices of graph $G = (V,E)$ and $m$  the number of edges of graph $G$.

\subsubsection{Finding twins and false twins}
The subroutine \emph{gettwins} can be found in appendix \ref{gettwins}. In line 2-10 two lists, \emph{nbs} and \emph{nbs2}, are created. Both lists contains $n$ sorted sublists. In the former list each sublist contains all the neighbours of a node and in the latter each sublist contains all the neighbours of a node and also the node itself. Finding all the neighbours only requires going though all the edges once, so the time complexity is $O(m)$. The sorting algorithm has time complexity $O(y \log y)$, $y=$max(len($x$)), $x \in $ \emph{nbs} < $O(n \log n)$, but since it is in a for loop of length $n$ the overall time complexity becomes at most $O(n^2 \log n)$.

In line 11-16 a dictionary \emph{falsetwins} is created where the keys are all the items from \emph{nbs} and the values are lists of the corresponding vertices. The worst case time complexity of checking if an item is in a dictionary is $O(k+l)$ where $k$ is the number of keys and $l$ the length of the key, which is a tuple. However, the average case is only $O(l)$ \cite{website:pythonTimeComplexity}.  $l\leq n$ and $k\leq n$, and this dictionary operation is in a for loop of length $n$, so the worst case time complexity of this is $O(n^3)$, but the average case is $O(n^2)$. In line 23 the items in the \emph{falsetwins} with only one vertex as value are removed. This is $O(n)$, since len() has $O(1)$.\cite{website:pythonTimeComplexity} The remaining values of the dictionary \emph{falsetwins} are now all the sets of false twins as defined in definition \ref{deffalsetwins}. A similar dictionary \emph{twins} is created where the keys are the items from \emph{nbs2}. The worst case time complexity of \textit{gettwins} is thus $O(n^3)$.

\subsubsection{Replacing the twins}
The subroutine \emph{preprocessing} can be found in appendix \ref{preprocessing}. In line 2 the twins and falsetwins are defined and if there are none, the program will terminate. If there are any twins, all edges with a head or tail in one of these twins will be deleted. The complexity of this is $O(mv) < O (nm)$, where $v$ is the number of vertices in (\emph{twins}+\emph{falsetwins}). Then in line 13-19 all vertices which are in twins or falsetwins will be removed. The complexity of this is $O(nv) < O(n^2)$. In line 20-26 a new colored vertex will be added for each (false) twin $(O(n))$ and finally the edges from these vertices will be added ($O(n^2 (n+m))$). The worst case time complexity of this algorithm is ($O(n^2 (n+m)$), but if there are only $v$ vertices which are \emph{twins} or \emph{falsetwins} and they have maximally $w$ neighbours, the complexity is $O(vw) (n+m)$.

%------------------------------------------------
\section{Proof of Correctness}
\begin{definition}
A \emph{colored} graph $G(E,V)$ is a graph where each vertex $v \in V$ has a color label.
\end{definition}
\begin{definition}\label{def:iso}
Let $G_1 = (V_1, E_1)$ and $G_2 = (V_2, E_2)$ be two undirected graphs. A function $f: V_1 \to V_2$ is called a \emph{graph isomorphism} if:
\begin{compactenum}[\upshape(a)]
\item $f$ is a bijection. \label{voorwaarde1}
\item $\forall a, b \in V_1, \{a, b\} \in E_1$ if and only if $\{f(a), f(b)\} \in E_2$.\label{voorwaarde2}
\end{compactenum}
When such a function exists, $G_1$ and $G_2$ are called \emph{isomorphic graphs}. 
\end{definition}
\begin{definition}\label{def:coloriso}
Let $G_1 = (V_1, E_1)$ and $G_2 = (V_2, E_2)$ be two undirected colored graphs. A function $f: V_1 \to V_2$ is called a \emph{colored graph isomorphism} if: 
\begin{compactenum}[\upshape(a)]
\item $f$ is a graph isomorphism.\label{voorwaarde4}
\item $\forall v \in V_1$: $color(v)$ = $color (f(v)$. \label{voorwaarde3}
\end{compactenum}
When such a function exists, $G_1$ and $G_2$ are called \emph{colored isomorphic graphs}. 
\end{definition}
\begin{theorem}
Let $G_1 = (V_{G1}, E_{G1})$ and $G_2 = (V_{G2}, E_{G2})$ be two undirected  graphs and let $H_1 = (V_{H1}, E_{H1})$ and $H_2 = (V_{H2}, E_{H2})$ be resp. the same graphs as $G_1$ and $G_2$ where vertices which are not (false) twins with any other vertex get color 0 and each (false) twin is replaced by a colored vertex such that: \begin{compactenum}[\upshape(a)]
\item Vertices in $H_1$ ($H_2$) have the same color $\iff$ The corresponding twinsets or false twinsets in $G_1$ ($G_2$) have the same size (number of vertices).
\item Two vertices in $H_1$ ($H_2$) cannot have the same color if one of them corresponds to a twinset in $G_1$ ($G_2$) and the other one corresponds to a false twinset in $G_1$ ($G_2$). They also cannot have color 0.
\end{compactenum}
Graph $G_1$ and $G_2$ are isomorphic $\iff$ Graph $H_1$ and $H_2$ are colored isomorphic.
\end{theorem}
\begin{proof}
%We shall define $f_i$: $G_i\to H_i$ as the transformation descriped above and   and $f_i^{-1}$: $H_i\to G_i$ as the inverse of that transformation. Note that $f_i^{-1}$ maps a vertex with color $\neq$ 0 to a set of twins or falsetwins. Also w.l.o.g we assume $f$ gives color (

$\Leftarrow$:  $\exists$ colored graph isomorphism $h: H_1 \to H_2$. If $v \in V_{H1}$ has color 0: we define $g(v) = h(v)$.
If $v \in V_{H1}$ has color $\neq$ 0, $\exists$ $ w \in V_{H2}$ such that color($v$) = color($w$) (since $H_1$ and $H_2$ are isomorphs). $v$ and $w$ either have resp. corresponding false twinsets \emph{ftw} $\subseteq V_{G1}$ and \emph{ftw2} $\subseteq V_{G2}$ or they have resp. corresponding twinsets \emph{tw} $\subseteq V_{G1}$ and \emph{tw2} $\subseteq V_{G2}$. In the first case $g$ maps each $x \in$ \emph{ftw} to a different $y \in$ \emph{ftw2}. Since they are false twins it does not matter which $x$ is mapped to which $y$. In the second case $g$ maps $x \in$ \emph{tw} to $y \in$ \emph{tw2} similarly.

We want to prove that $g: G_1 \to G_2$ is a graph isomorphism. $V_{H1}$ and $V_{H2}$ have the same size and each $v_1 \in V_{H1}$ corresponds to a $v_2 \in V_{H1}$ with the same color. The colors of $v_1$ and $v_2$ defines the sizes of the corresponding (false) twinsets in resp. $G_{H1}$ and $G_{H2}$, thus $G_{H1}$ and $G_{H2}$ also have the same size. Since $g$ is defined such that it is an injection and $V_{G1}$ and $V_{G2}$ have the same size, $g$ is a bijection \cite{website:proofwiki}. Thus condition \ref{voorwaarde1} of definition \ref{def:iso} is satisfied.

Define a module $M_i \subseteq V_{G1}: \{v \in V_{G1}$ $|$ corresponding vertex in $H_1$ has color $i \}$. $\forall a, b \in V_{G1}$, $a \in M_i$ and $b$ $\in$ $M_j$ for some $i,j$: 
\begin{compactenum}
\item if $i = j$: $\{a,b\} \in E_{G1} \iff \{c,d\} \in E_{H1}$. $c,d$ are the corresponding vertices of resp. $a,b$. $\{c,d\} \in E_{H1} \iff \{h(c),h(d)\} \in E_{H2} \iff \{g(a),g(b)\} \in E_{G2}$, because $g(a)$ and $g(b)$ are elements of the corresponding (false) twinset of resp. $c$ and $d$ and all those elements have the same neighbours. 
\item if $i \neq j$: If $M_i$ is a false twin, $\{a,b\} \notin E_{G1} $ and $\{g(a),g(b)\} \in E_{G2}$. If $M_j$ is a twin $\{a,b\} \in E_{G1} $ and $\{g(a),g(b)\}\in E_{G2}$. If $i = 0$, the loop consists in $G_1 \iff$ it exists in $H_1 \iff$ it exists in $H_2 \iff$ it exists in $G_2$.
\end{compactenum} 
$\forall a, b \in V_{G1}, \{a, b\} \in E_{G1} \iff \{g(a), g(b)\} \in E_{G2}$.
Thus condition \ref{voorwaarde2} of definition \ref{def:iso} is satisfied and $g$ is a graph isomorphism.

$\Rightarrow$: $\exists$ graph isomorphism $g: G_1 \to G_2$. $\forall M_i (G_1) \subseteq V_{G1}, M_i (G_2) \subseteq V_{G2}$. The function $h$ maps the corresponding vertex of $M_i (G_1)$ in $H_1$ (with color i) to the corresponding vertex of $M_i (G_2)$ in $H_2$ (with color i).

We want to prove that $h: H_1 \to H_2$ is a colored graph isomorphism. It is defined such that it is color preserving, so condition \ref{voorwaarde3} of definition \ref{def:coloriso} is satisfied. For each $M_i (G_1) \subseteq V_{G1}$ $\exists$ $M_i (G_2) \subseteq V_{G2}$ with the same size and each $M_i$ in $G$ is a single vertex in $H$. Thus $V_{H1}$ and $V_{H2}$ have the same size (since $V_{G1}$ and $V_{G2}$ have the same size). Since $G_1$ and $G_2$ are isomorphic, they contain the same modules $M_i$. Thus $h$ can map each corresponding vertex of $M_i (G1)$ to a unique corresponding vertex of $M_i (G2)$. So $h$ is an injection and $V_{H1}$ and $V_{H2}$ have the same size, thus $h$ is also a bijection \cite{website:proofwiki}.

$\forall a, b \in V_{H1}$: color($a$) = i, color($b$) = j, $a$ corresponds to $M_i (G_1)$ and $b$ corresponds to $M_j (G_1)$. If $i=j: M_i$ and $M_j$ can either be the same set of (false) twins or different sets of equal size and both twins or both false twins.  $ \{a, b\} \in E_{H1} \iff \{c,d\} \in E_{G1}, c \in  M_i, d \in  M_j \iff \{g(c),g(d)\} \in E_{G2} \iff \{h(a),h(b)\} \in E_{H2}$. Because $h(a)$ and $h(b)$ corresponds to resp. $g(c)$ and $g(d)$. By definition \ref{def:iso}, $h$ is graph isomorphism. Condition \ref{voorwaarde4} of definition \ref{def:coloriso} is satisfied and $h$ is a colored graph isomorphism.

\end{proof}

%------------------------------------------------
\section{Results}

The results of the experiment are shown in table \ref{table:results}. Seven sets of graphs were tested for isomorphisms and in three cases the computation time was significantly shorter. In the other cases, the preprocessing time is relatively short compared to the total computing time. This is because the low complexity of \emph{gettwins} and the fact that \emph{preprocessing} does nothing when there are no (false) twins. The difference in computation time is not related to the preprocessing time, as the preprocessing time is not equal to the difference in computation times. This is likely due to the circumstances of both executions of the algorithms not occurring under the exact same circumstances, as it was impossible to ensure the conditions were the same.

Due to the large differences in computation times relative to the preprocessing time, only significant differences in computation time will be used as indication of increase or decrease in speed.

The preprocessing times for single graphs with $n$ nodes can be found in table \ref{table:results2}. These graphs contain no (false) twins, so these times are  only of the subroutine \emph{gettwins}. This is the extra computation time in case the preprocessing does nothing.

%------------------------------------------------
\section{Conclusions}

In three of the seven test cases the computation time increased significantly. In the other four there was no significant difference in the computation times.

Thus, it can be concluded that this preprocessing method reduces the computation time needed to find isomorphisms between graphs that have (false) twins.

%----------------------------------------------------------------------------------------
%	REFERENCE LIST
%----------------------------------------------------------------------------------------

\printbibliography
%bibliography{references}

%----------------------------------------------------------------------------------------

\end{multicols}
\begin{appendices}

\section{Python code}
\lstset{language=Python}
\begin{lstlisting}[label={gettwins}]
def gettwins(g):
	nbs = []
	nbs2 = []
	for vertex in g.V():
		nb = vertex.nbs()[:]
		nb.sort(key=lambda x: x._label)
		nbs.append(tuple(nb))
		nb.append(vertex)
		nb.sort(key=lambda x: x._label)
		nbs2.append(tuple(nb))
	falsetwins = {}
	for i, item in enumerate(nbs):
		if item in falsetwins.keys():
			falsetwins[item].append(g.V()[i])
		else:
			falsetwins[item] = [g.V()[i]]
	twins = {}
	for i, item in enumerate(nbs2):
		if item in twins.keys():
			twins[item].append(g.V()[i])
		else:
			twins[item] = [g.V()[i]]
	falsetwins = {k: v for k, v in falsetwins.items() if len(v) > 1}
	twins = {k: v for k, v in twins.items() if len(v) > 1}
	return list(falsetwins.values()), list(twins.values()), list(falsetwins.keys()), list(twins.keys())   
\end{lstlisting}

\begin{lstlisting}[label={preprocessing}]
def preprocessing(g):
	falsetwins, twins, falsetwinsN, twinsN = gettwins(g)
	lftwins = len(falsetwins)
	ltwins = len(twins)
	if lftwins != 0 or ltwins != 0:
		deledges = []
		for i, e in enumerate(g._E):
			if any(e._tail in ftwin for ftwin in falsetwins) or any(e._head in ftwin for ftwin in falsetwins) or any(e._tail in twin for twin in twins) or any(e._head in twin for twin in twins):
		deledges.append(i)
		deledges.sort(reverse=True)
		for i in deledges:
			g._E.pop(i)
		delnodes = []
		for i, V in enumerate(g._V):
			if any(V in ftwin for ftwin in falsetwins) or any(V in twin for twin in twins):
				delnodes.append(i)
		delnodes.sort(reverse=True)
		for i in delnodes:
			g._V.pop(i)
		combined = falsetwinsN + twinsN
		for i, twin in enumerate(falsetwins):
			g.addvertexobject(twin[0], True, len(twin))
			twin[0].colornum = -twin[0].twinsize
		for i, twin in enumerate(twins):
			g.addvertexobject(twin[0], True, len(twin))
			twin[0].colornum = twin[0].twinsize
		for i, twin in enumerate(falsetwins + twins):
			for j in combined[i]:
				if j in g.V() and twin[0] != j and not g.findedge(twin[0],k)
					g.addedge(twin[0], j)
		g._V.sort(key=lambda x: x._label)
	return g, lftwins, ltwins
\end{lstlisting}

\section{Tables with computation times}
\begin{table}[h]\centering
\begin{tabular}{c|C{0.3\textwidth}|C{0.3\textwidth}|c}
Graph & Computation time (without preprocessing) & Computation time (with preprocessing) & Preprocessing time \\
\hline
cographs1 & 6.6017 sec & 0.0048 sec & 0.0027 sec \\
hugecographs & >7 hours & 29.8713 sec & 0.0214 sec \\
cycles175 & 13.7522 sec & 13.9265 sec & 0.0238 sec\\
bigtrees3 & 0.9652 sec & 0.2700 sec & 0.0323 sec\\
cubes6 & 0.5967 sec & 0.5877 sec & 0.0072 sec\\
products72 & 0.6083 sec & 0.6353 sec & 0.0123 sec\\
torus72 & 3.5590 sec & 3.6059 sec & 0.0106 sec\\
\end{tabular}
\caption{Table showing the result from the experiment. For each set of graphs, it shows the computation time without and with preprocessing time, as well as the time it took to perform the preprocessing.}
\label{table:results}
\end{table}
\begin{table}[h]\centering

\begin{tabular}{c|c}
n & time \\
\hline
34 & 0.0017\\
66 & 0.0032\\
130 & 0.0071\\
258 & 0.0161\\
514 & 0.0344\\
1026 & 0.0805\\
2050 & 0.2019\\
4098 & 0.5874\\
8194 & 2.7102

\end{tabular}
\caption{Table showing various preprocessing times, for graphs with $n$ nodes and no (false) twins.}
\label{table:results2}
\end{table}
\end{appendices}

\end{document}

### 2016-03-16 paper/paper.tex
Additions to paper, including todo's for every sections and some more comment to clarify stuff

@@
 \usepackage[sc]{mathpazo} % Use the Palatino font
+% Output encoding
 \usepackage[T1]{fontenc} % Use 8-bit encoding that has 256 glyphs
+% Input encoding
+\usepackage[utf8]{inputenc} % UTF-8 character encoding stuff
 \linespread{1.05} % Line spacing - Palatino needs more space between lines
@@
 \bibliography{references.bib}
+
+% Appendices
+\usepackage[toc,page]{appendix} % appendix
 
@@
 \section{Scope \& Objectives}
+The aim of this research paper is to calculate and compare the worst-case time complexity of both color refinement algorithms and verifying them using computational results.
+These results will be used to examine the differences in computation time between the algorithms for different types of graphs.
+
+This paper will describe both algorithms, give the calculations for the asymptotic worst-case time complexity, as well as an examination of the computational results.
+%------------------------------------------------
+\section{Methods}
+%% TODO %%
+% Describe experiment, setup, and method of testing
+%%
 
 %------------------------------------------------
-\section{Methods}
-
-%------------------------------------------------
-\section{Algorithms} %Descriptions + time complexity
+\section{Algorithms}
+%% TODO %%
+% Describe algorithms, show concise calculations of time complexity
+%%
 
@@
 \section{Results}
+%% TODO %%
+% Give and describe results
+%%
+
+%------------------------------------------------
+\section{Discussion} %Change section name?
+%% TODO %%
+% Describe the effectiveness of the algorithms on various types of graphs, and examine the differences
+%%
 
@@
 \section{Conclusions}
+%% TODO %%
+% Draw conlusions from both the results and discussion
+%%
 
@@
 %----------------------------------------------------------------------------------------
+%% TODO %%
+% Put all the things that should be in the paper, but don't fit the main part here
+%%
 
+% Everything EXCEPT multicols can be removed from the appendices if none are added.
+\end{multicols}
+
+\begin{appendices}
+\end{appendices}
 %----------------------------------------------------------------------------------------
 
-\end{multicols}
-%\begin{appendices}
-
-%----------------------------------------------------------------------------------------
-
-%\end{appendices}
 \end{document}

### 2016-03-23 paper/paper.tex
stuff

@@
 
-\noindent Test
+\noindent This will be the abstract of the paper
 
@@
 \section{Introduction}
+\label{intro}
 This paper will compare the time complexity of two color refinement algorithms and confirm them with computational results.
@@
 \section{Scope \& Objectives}
+\label{scope}
 The aim of this research paper is to calculate and compare the worst-case time complexity of both color refinement algorithms and verifying them using computational results.
@@
 \section{Methods}
+\label{methods}
 %% TODO %%
@@
 \section{Algorithms}
+\label{algs}
 %% TODO %%
@@
 \section{Results}
+\label{results}
 %% TODO %%
@@
 \section{Discussion} %Change section name?
+\label{disc}
 %% TODO %%
@@
 \section{Conclusions}
+\label{concl}
 %% TODO %%
@@
 %----------------------------------------------------------------------------------------
-
+\label{references}
 \printbibliography
@@
 \begin{appendices}
+\label{appendices}
 \end{appendices}

### 2016-03-23 colorRefinement.py
added two large examples, and increasing recursionlimit. this might cause segmentation faults

@@ -1,3 +1,5 @@
+import sys
+
 __author__ = 'Tim (& [Jeroen])'
 
 import copy
@@ -435,7 +437,10 @@ def componentpreproc(graphlisturl):
 
 start_time = time.clock()
 
-compare("GI_TestInstancesWeek1/products72.grl", False)  # #aut for product72
+sys.setrecursionlimit(10000)
+
+compare("GI_TestInstancesWeek1/colorref_largeexample_6_960.grl", True)  # GI for large 4 1026
+# compare("GI_TestInstancesWeek1/products72.grl", False)  # #aut for product72
 # compare("GI_TestInstancesWeek1/torus72.grl", False)  # #aut for torus72
 # compare("GI_TestInstancesWeek1/cubes6.grl", True)  # GI for cubes6
 # compare("GI_TestInstancesWeek1/bigtrees3.grl", True)  # GI for bigtrees3

### 2016-03-23 experiment/algorithm.py
(slow) color refinement

@@ -1,0 +1,127 @@
+import copy
+
+from experiment import graphIO as io
+from experiment.basicgraphs import *
+
+
+def initial_coloring(g: Graph) -> (Graph, dict):
+    color_dict = {}
+    for v in g.vertices():
+        v.colornum = v.degree()
+    for v in g.vertices():
+        if v.colornum not in color_dict:
+            color_dict[v.colornum] = (v.degree(), create_nbs_dict(v))
+    return g, color_dict
+
+
+def create_nbs_dict(v: Vertex, update: bool = False) -> {}:
+    """
+    Creates a dictionary holding the colors of the neighbourhood of vertex v, and their number of appearance.
+    :param v:
+    :param update:
+    :return:
+    """
+    nbs_dict = {}
+    for nb in v.nbs():
+        c = nb.colornum
+        if update and hasattr(nb, 'new_colornum'):
+            c = nb.new_colornum
+        if c in nbs_dict:
+            nbs_dict[c] += 1
+        else:
+            nbs_dict[c] = 1
+    return nbs_dict
+
+
+def new_color(color_dict: {}):
+    c = 0
+    while c in color_dict:
+        c += 1
+    return c
+
+
+def color_refine(g: Graph, color_dict: {}) -> Graph:
+    # color_dict{key, value} = {colornum, (degree, nbs_dict)}
+    changed = True
+    while changed:
+        changed = False
+        for v in g.vertices():
+            nbs_dict = create_nbs_dict(v)
+            if color_dict[v.colornum][1] == nbs_dict:
+                # dont update, as the nbs are equivalent
+                pass
+            else:
+                existing_config = False
+                for c, (d, n) in color_dict.items():
+                    if d == v.degree() and n == nbs_dict:
+                        # already assigned this configuration a new color
+                        v.new_colornum = c
+                        existing_config = True
+                        break
+                if not existing_config:
+                    # This configuration is not yet known, assign a new color
+                    c = new_color(color_dict)
+                    v.new_colornum = c
+                    color_dict[c] = (v.degree(), nbs_dict)
+
+        for v in g.vertices():
+            # update colors
+            if hasattr(v, 'new_colornum'):
+                v.colornum = v.new_colornum
+                del v.new_colornum
+                changed = True
+            # update color_dict
[…]

### 2016-03-23 experiment/basicgraphs.py

@@ -1,0 +1,259 @@
+"""
+This is a module for working with *undirected* graphs (simple graphs or multigraphs).
+
+It contains three classes: vertex, edge and graph. 
+
+The interface of these classes is extensive and allows programming all kinds of graph algorithms.
+
+However, the data structure used is quite basic and inefficient: a graph object stores only a vertex list and an edge
+list, and methods such as adjacency testing / finding neighbors of a vertex require going through the entire edge list!
+"""
+# version: 29-01-2015, Paul Bonsma
+
+
+unsafe = False
+
+
+# Set to True for faster, but unsafe listing of all vertices and edges.
+
+class GraphError(Exception):
+    def __init__(self, message):
+        self.mess = message
+
+    def __str__(self):
+        return self.mess
+
+
+class Vertex:
+    """
+    Vertex objects have an attribute <_graph> pointing to the graph they are part of,
+    and an attribute <_label> which can be anything: it is not used for any methods,
+    except for __repr__.
+    """
+
+    def __init__(self, graph, label=0):
+        """
+        Creates a vertex, part of <graph>, with optional label <label>.
+        (Labels of different vertices may be chosen the same; this does
+        not influence correctness of the methods, but will make the string
+        representation of the graph ambiguous.)
+        """
+        self._graph = graph
+        self._label = label
+        self._incidence_list = []
+        self._degree = 0
+
+    def __repr__(self):
+        if hasattr(self, "colornum"):
+            return"(" + str(self._label) + ", c=" + str(self.colornum) + ")"
+        else:
+            return str(self._label)
+
+    def adjacent(self, other):
+        """
+        Returns True iff vertex <self> is adjacent to <other> vertex.
+        """
+        return self._graph.adjacent(self, other)
+
+    def incidence_list(self):
+        """
+        Returns the list of edges incident with vertex <self>.
+        """
+        return self._incidence_list
+
+    def nbs(self):
+        """
+        Returns the list of neighbors of vertex <self>.
+        In case of parallel edges: duplicates are not removed from this list!
+        """
+        nbl = []
+        for e in self.incidence_list():
+            nbl.append(e.other_end(self))
+        return nbl
+
+    def degree(self) -> int:
+        """
+        Returns the degree of vertex <self>.
+        """
+        return self._degree
+
+    def add_nb(self, e):
+        self._incidence_list.append(e)
+        self._degree += 1
+
+
+
+
+
+class Edge:
+    """
[…]

### 2016-03-23 experiment/graphIO.py

@@ -1,0 +1,259 @@
+"""
+Includes functions for reading and writing graphs, in a very simple readable format.
+ loadgraph:		reads from files
+ inputgraph:	reads from terminal input / stdin
+ savegraph:		writes to files
+ printgraph:	writes to terminal / stdout
+ writeDOT:		writes in .dot format; can be used for visualization.
+ 
+The other functions are internal, to implement the above functions.  
+
+The graph objects returned by loadgraph and inputgraph are by default constructed using the <graph> class in the
+module basicgraphs.py, but by using an optional argument you can use your own graph class (provided that it supports
+the same methods/interface).
+
+This module also supports edge weighted graphs: edges should/will have an (integer) attribute <weight>. 
+"""
+# Version: 30-01-2015, Paul Bonsma
+# updated 30-01-2015: writeDOT also writes color information for edges.
+# updated 2-2-2015: writeDOT can also write directed graphs.
+# updated 5-2-2015: no black fill color used, when more than numcolors**2 vertices.
+from experiment import basicgraphs
+
+defaultcolorscheme = "paired12"
+numcolors = 12
+
+
+# defaultcolorscheme="piyg11"
+# numcolors=11
+
+# defaultcolorscheme="spectral11"
+# numcolors=11
+
+def readgraph(graphclass, readline):
+    """
+    For internal use.
+    """
+    options = []
+    while True:
+        try:
+            S = readline()
+            n = int(S)
+            G = graphclass(n)
+            break
+        except ValueError:
+            if len(S) > 0 and S[-1] == '\n':
+                options.append(S[:-1])
+            else:
+                options.append(S)
+    S = readline()
+    edgelist = []
+    try:
+        while True:
+            comma = S.find(',')
+            if ':' in S:
+                colon = S.find(':')
+                edgelist.append((int(S[:comma]), int(S[comma + 1:colon]), int(S[colon + 1:])))
+            else:
+                edgelist.append((int(S[:comma]), int(S[comma + 1:]), None))
+            S = readline()
+    except Exception:
+        pass
+    for edge in edgelist:
+        # print("Adding edge (%d,%d)"%(edge[0],edge[1]))
+        e = G.add_edge(G[edge[0]], G[edge[1]])
+        if edge[2] is not None:
+            e.weight = edge[2]
+    if S != '' and S[0] == '-':
+        return G, options, True
+    else:
+        return G, options, False
+
+
+def readgraphlist(graphclass, readline):
+    """
[…]

### 2016-03-23 paper/paper.tex
methods

@@
 \label{methods}
-%% TODO %%
-% Describe experiment, setup, and method of testing
-%%
+%% WIP %%
+The time complexity for both algorithms was calculated by hand.
+To verify the results, a computational experiment was performed.
+In this experiment, both algorithms were tested with various graphs from the test set.
+
+The testing was all done on the same machine, under the same circumstances. Every test was performed at least three times in order to prevent outliers caused by external events.
+
+\subsection{Testing}
+\label{methods:testing}
+In the experiment, both algorithms were tested with several sizes of the same type of graph, for various types of graphs. 
+% TODO Insert method of testing here
+Only the algorithms themselves were timed, initialization and other parts of the code were not included.
 

### 2016-04-05 experiment/algorithm.py
Doubly linked list and some small stuff

@@ -1,6 +1,6 @@
 import copy
 
-from experiment.basicgraphs import *
+from experiment.structures.basicgraphs import *
 
 
 def initial_coloring(g: Graph):
@@ -115,7 +115,6 @@ def branching(g: Graph):
                 h.color_dict[branch_color] = (bv.degree(), create_nbs_dict(bv))
                 color_refine(h)
                 return isomorph_test(h)
-    print('fuck')
 
 
 def isomorph_test(g: Graph):

### 2016-04-05 experiment/io/graphIO.py

@@ -18,7 +18,7 @@ This module also supports edge weighted graphs: edges should/will have an (integ
 # updated 30-01-2015: writeDOT also writes color information for edges.
 # updated 2-2-2015: writeDOT can also write directed graphs.
 # updated 5-2-2015: no black fill color used, when more than numcolors**2 vertices.
-from experiment import basicgraphs
+from experiment.structures import basicgraphs
 
 defaultcolorscheme = "paired12"
 numcolors = 12

### 2016-04-05 experiment/main.py

@@ -1,14 +1,19 @@
+import sys
 from enum import Enum
 
-from experiment import graphIO as io
 from experiment.algorithm import *
+from experiment.io import graphIO as io
 
 # graph_file = '../graphs/colorref_smallexample_6_15.grl'
-# graph_file = '../graphs/torus24.grl'
+# graph_file = '../graphs/colorref_largeexample_4_1026.grl'
+# graph_file = '../graphs/torus144.grl'
 # graph_file = '../graphs/trees90.grl'
 graph_file = '../graphs/products72.grl'
+
+
 # graph_file = '../graphs/cographs1.grl' # TODO: goes out of range, colors aren't freed or something?
-# graph_file = '../graphs/bigtrees1.grl'
+# raph_file = '../graphs/bigtrees1.grl'
+# graph_file = '../graphs/wheelstar12.grl'
 
 
 def iso_check(g: Graph, h: Graph):
@@ -36,9 +41,8 @@ def main():
     graphs = io.loadgraph(filename=graph_file, readlist=True)[0]
     isomorphisms = []
     number_of_graphs = len(graphs)
-    if number_of_graphs < 2:
-        print("Only 1 graph in graphlist.")
-    else:
+    print("Number of graphs: " + str(number_of_graphs))
+    if number_of_graphs > 1:
         if iso_check(graphs[0], graphs[1]) == Result.ISOMORHP:
             isomorphisms.append([0, 1])
         else:
@@ -47,17 +51,26 @@ def main():
 
         if number_of_graphs > 2:
             for i in range(2, number_of_graphs):
-                for j in isomorphisms:
-                    if iso_check(graphs[i], graphs[j[0]]) == Result.ISOMORHP:
-                        isomorphisms[j[0]].append(i)
+                j = 0
+                iso = False
+                while j < len(isomorphisms):
+                    iso = Result.ISOMORHP == iso_check(graphs[i], graphs[isomorphisms[j][0]])
+                    if iso:
+                        isomorphisms[j].append(i)
                         break
-                isomorphisms.append([i])
+                    j += 1
+                if not iso:
+                    isomorphisms.append([i])
 
-    # prints isomorphisms
     for i in isomorphisms:
         if len(i) > 1:
             print(i)
 
 
+def refine_only():
+    graph = graphs = io.loadgraph(filename=graph_file, readlist=True)[0][0]
+
+
 if __name__ == '__main__':
+    sys.setrecursionlimit(10000)
     main()

### 2016-04-05 experiment/structures/basicgraphs.py

@@ -14,8 +14,6 @@ list, and methods such as adjacency testing / finding neighbors of a vertex requ
 # Improved: 28-03-2016, Rick Fontein
 
 
-
-
 class GraphError(Exception):
     def __init__(self, message):
         self.mess = message

### 2016-04-05 experiment/structures/doublylinkedlist.py

@@ -1,0 +1,67 @@
+class _DoublyNode:
+    def __init__(self, data, prev, next_):
+        self.data = data
+        self.prev = prev
+        self.next = next_
+        pass
+
+    pass
+
+
+class DoublyLinkedList:
+    def __init__(self):
+        self._head = None
+        self._tail = None
+        pass
+
+    def append(self, data):
+        new_node = _DoublyNode(data, None, None)
+        if self._head is None:
+            # empty list
+            self._head = self._tail = new_node
+        else:
+            # link nodes together
+            self._tail.next = new_node
+            new_node.prev = self._tail
+            # update tail
+            self._tail = new_node
+        pass
+
+    def remove(self, data, all_instances: bool = False):
+        pointer = self._head
+        p = None
+        while pointer is not None:
+            if pointer.data == data:
+                # first element is head
+                if pointer.prev is None:
+                    self._head = pointer.next
+                    self._head.prev = None
+                    p = pointer.next
+                    pointer.next = None
+                elif pointer.next is None:
+                    self._tail = pointer.prev
+                    self._tail.next = None
+                    pointer.prev = None
+                else:
+                    pointer.prev.next = pointer.next
+                    pointer.next.prev = pointer.prev
+                if not all_instances:
+                    break
+            if p is not None:
+                pointer = p
+                p = None
+            else:
+                pointer = pointer.next
+        pass
+
+    def __repr__(self):
+        pointer = self._head
+        if pointer is None:
+            return "[]"
+        r = "[" + str(pointer.data)
+        pointer = pointer.next
+        while pointer is not None:
+            r += ", "
+            r += str(pointer.data)
+            pointer = pointer.next
+        return r + "]"

### 2016-04-05 experiment/test.py
graphs

@@ -28,6 +28,10 @@ def main():
     print(min(Timer(t5).repeat(3, 1)))
     print(min(Timer(t6).repeat(3, 1)))
     print(min(Timer(t7).repeat(3, 1)))
+    print(min(Timer(t8).repeat(3, 1)))
+    print(min(Timer(t9).repeat(3, 1)))
+    print(min(Timer(t10).repeat(3, 1)))
+    print(min(Timer(t11).repeat(3, 1)))
 
 
 def t0():

### 2016-04-05 graphs.m
added the matlab file for graphs

@@ -1,0 +1,95 @@
+%% Data
+% Full Size Range
+size = [ 5 10 20 40 80 160 320 640 1280 2560 5120 10240 ];
+
+% Basic Color Refinement
+% Measurement 1
+sizeb1 = [ 5 10 20 40 80 160 320 640 1280 2560 5120 ];
+timeb1 = [ 0.000447894 0.001738771 0.007223303 0.031326249 0.125898654 0.604074251 3.213163267 19.26075537 127.5672593 907.4070568 5736.549588 ];
+% Slides w/out adjacency
+sizeb2 = [ 160 320 640 ];
+timeb2 = [ 7 54 427 ];
+% Slides w/ adjacency
+sizeb3 = [ 160 320 640 1280 ];
+timeb3 = [ 0.3 2 17 125 ];
+
+% Fast Color Refinement
+
+% Slides w/ adjacency
+sizef2 = [ 160 320 640 1280 2560 5120 10240 ];
+timef2 = [ 0.04 0.09 0.13 0.3 1 3 12 ];
+
+%% Computation Time Subplot
+
+%subplot(2,1,1);
+
+% Basic Measurement 1
+p = semilogy(sizeb1, timeb1);
+p.Color = 'blue';
+p.Marker = '.';
+p.MarkerSize = 8;
+
+hold on
+title('Computation Time of Basic Color Refinement for Threepaths')
+xlabel('Size \it (# vertices per path)')
+ylabel('Time \it (s)')
+xlabels = [ 5 320 640 1280 2560 5120 10240 ];
+set(gca, 'Xtick', xlabels);
+set(gca, 'XtickLabel', xlabels);
+
+% Basic Slides w/out Adjacency
+p = semilogy(sizeb2, timeb2);
+p.Color = 'magenta';
+p.Marker = 'o';
+p.MarkerSize = 6;
+
+% Basic Slides w/ Adjacency
+p = semilogy(sizeb3, timeb3);
+p.Color = 'green';
+p.Marker = 'o';
+p.MarkerSize = 6;
+
+% Fast Slides w/ Adjacency
+p = semilogy(sizef2, timef2);
+p.Color = 'cyan';
+p.Marker = 'o';
+p.MarkerSize = 6;
+
+% x^2
+x = 0:5:10240;
+p = semilogy(x, x.^2);
+p.Color = 'red';
+
+% x^3
+p = semilogy(x, x.^3);
+p.Color = 'black';
+
[…]

### 2016-04-05 paper/paper.tex
added template table for results, fixed month of reference
algorithms added for Huub, started on checking grammar

@@
 \usepackage[toc,page]{appendix} % appendix
+
+% Additional column type
+\usepackage{array}
+\newcolumntype{C}[1]{>{\centering\arraybackslash}p{#1}}
 
@@
 %%
+For the paper, two different color refinement methods were implemented.
+The basic one gives all vertices in a graph with the same color but differently colored neighborhoods a separate color.
+This is done iteratively until the graph is stably colored.
+A coloring $\alpha$ of $G$ is stable if for all $u, v \in V(G)$ it holds that: if $\alpha(u) = \alpha(v)$, then $u$ and $v$ have identically colored neighborhoods.
+The problem with this method is that all vertices are checked each iteration.
+With fast color refinement the colors that lead to refinements are tracked.
+These colors are maintained in a queue.
+If a color class is being refined, the class is split into two different classes, with one class keeping the old color.
+The algorithm terminates with a stable color if no color class needs refinement in other words the queue is empty.
+The initial coloring is done in a separate algorithm before the color refinement algorithms take place and is therefore the same for both the fast and basic implementation.
+Computing time of our algorithms are increasing by iterations over the vertices in a graph.
+To calculate the time complexity of the algorithms, the number of and the places of the iterations over the vertices are analyzed.
+In the basic implementation are several places where iterations over vertices take place.
+The code contains one main loop that takes worst case N iterations.
+So the time complexity so far is $O(\boldsymbol{N})$ In this main loop an iteration takes place over all the vertices in the graph.
+Time complexity is $O(N \cdot \boldsymbol{N})$ now.
+Each vertex is checked if it needs an update.
+If so, a new color dictionary is made and a new loop over the dictionary takes place.
+The time complexity is $O(N \cdot N \cdot \boldsymbol{N})$ for making the dictionary and gets $O((N \cdot N)(N+\boldsymbol{N}))$ for iterating over the dictionary.
+For each item in the dictionary is checked if the configuration already is assigned a new color.
+If the the configuration is not know yet, a new color is assigned.
+In worst case this new color operation costs again N.
+The time complexity is $O((N \cdot N)(N+N+\boldsymbol{N}))$ so far.
+After the first loop over the vertices in the main loop, again the algorithm iterates over all the vertices to color all the vertices correctly in the main dictionary.
+The time complexity worst case at the end is $O((N \cdot (N+\boldsymbol{N})(N+N+N)) = O(N^3)$
 
@@
 %%
+% Table for results can be found at appendices
 
@@
 \begin{appendices}
-\label{appendices}
+\label{app}
+
+%------------------------------------------------
+\section{Tables} %Change section name
+\label{app:tab}
+%% TODO %%
+% Fill table with actual values
+%%
+%% NOTE %%
+% All current values are fictional values for testing purposes only
+%%
+\begin{table}[h]\centering
+\label{app:tab:0}
+\begin{tabular}{|l|r@{.}l|r@{.}l|}
+\hline
+Graph	
+	& \multicolumn{2}{l|}{Algorithm A \emph{(s)}}
+	& \multicolumn{2}{l|}{Algorithm B \emph{(s)}} \\
+\hline
+graph\_{}type1\_{}16
+	& 0	& 001
+	& 0	& 005 \\
+graph\_{}type1\_{}64
+	& 0	& 1	
+	& 0	& 1 \\
+graph\_{}type1\_{}256
+	& 10	& 0	
+	& 2		& 0\\
+graph\_{}type1\_{}1024
+	& 1000	& 0 
+	& 40	& 0 \\
+graph\_{}type1\_{}4096
+	& 100000	& 0
+	& 800		& 0 \\
+\hline
+\end{tabular}
+\caption{This is a table}
+\end{table}
+%------------------------------------------------
+
 \end{appendices}

### 2016-04-05 colorRefinement.py
testing correctness, it is mostly correct. preprocessing might be slightly unpredictable

@@ -439,11 +439,11 @@ start_time = time.clock()
 
 sys.setrecursionlimit(10000)
 
-compare("GI_TestInstancesWeek1/colorref_largeexample_6_960.grl", True)  # GI for large 4 1026
-# compare("GI_TestInstancesWeek1/products72.grl", False)  # #aut for product72
-# compare("GI_TestInstancesWeek1/torus72.grl", False)  # #aut for torus72
-# compare("GI_TestInstancesWeek1/cubes6.grl", True)  # GI for cubes6
-# compare("GI_TestInstancesWeek1/bigtrees3.grl", True)  # GI for bigtrees3
+comparepreproc("GI_TestInstancesWeek1/torus24.grl", False)  # GI for large 4 1026
+comparepreproc("GI_TestInstancesWeek1/trees90.grl", False)  # #aut for torus72
+comparepreproc("GI_TestInstancesWeek1/products72.grl", False)  # #aut for product72
+comparepreproc("GI_TestInstancesWeek1/cographs1.grl", False)  # GI for cubes6
+compare("GI_TestInstancesWeek1/bigtrees1.grl", True)  # GI for bigtrees3
 #
 # comparepreproc("GI_TestInstancesWeek1/cographs1.grl")  # GI for cographs1 with preprocessing
 # comparepreproc("GI_TestInstancesWeek1/products72.grl")  # GI for cographs1 with preprocessing

### 2016-04-05 paper/paper.tex
changes to grammar stuff

@@
 The problem with this method is that all vertices are checked each iteration.
-With fast color refinement the colors that lead to refinements are tracked.
-These colors are maintained in a queue.
+
+With fast color refinement the colors that lead to refinements are tracked.(TODO: make better explanation) These colors are maintained in a queue.
 If a color class is being refined, the class is split into two different classes, with one class keeping the old color.
-The algorithm terminates with a stable color if no color class needs refinement in other words the queue is empty.
+The algorithm terminates with a stable coloring if no color class needs refinement, in other words the queue is empty.
+
 The initial coloring is done in a separate algorithm before the color refinement algorithms take place and is therefore the same for both the fast and basic implementation.
-Computing time of our algorithms are increasing by iterations over the vertices in a graph.
-To calculate the time complexity of the algorithms, the number of and the places of the iterations over the vertices are analyzed.
-In the basic implementation are several places where iterations over vertices take place.
-The code contains one main loop that takes worst case N iterations.
-So the time complexity so far is $O(\boldsymbol{N})$ In this main loop an iteration takes place over all the vertices in the graph.
-Time complexity is $O(N \cdot \boldsymbol{N})$ now.
-Each vertex is checked if it needs an update.
+
+The computation time of our algorithms increases for iterations over the vertices in a graph.
+To calculate the time complexity of the algorithms, the number and places of the iterations over the vertices are analyzed.
+
+The basic implementation has several places where iterations over vertices take place.
+The code contains one main loop that has a worst case of N iterations.
+Thereby the time complexity so far is $O(\boldsymbol{N})$, where $N = V(G)$.
+In this main loop an iteration takes place over all the vertices in the graph.
+The time complexity is now $O(N \cdot \boldsymbol{N})$.
+Each vertex is checked to see if it needs an update.
 If so, a new color dictionary is made and a new loop over the dictionary takes place.
-The time complexity is $O(N \cdot N \cdot \boldsymbol{N})$ for making the dictionary and gets $O((N \cdot N)(N+\boldsymbol{N}))$ for iterating over the dictionary.
-For each item in the dictionary is checked if the configuration already is assigned a new color.
-If the the configuration is not know yet, a new color is assigned.
-In worst case this new color operation costs again N.
+This makes the time complexity is $O(N \cdot N \cdot \boldsymbol{N})$ for making the dictionary and $O((N \cdot N)(N+\boldsymbol{N}))$ for iterating over the dictionary.
+Every color class in the dictionary is checked to see if the configuration has already been assigned a new color.
+If the the configuration is not yet know, a new color is assigned.
+In worst case this new color operation costs N.
 The time complexity is $O((N \cdot N)(N+N+\boldsymbol{N}))$ so far.
-After the first loop over the vertices in the main loop, again the algorithm iterates over all the vertices to color all the vertices correctly in the main dictionary.
-The time complexity worst case at the end is $O((N \cdot (N+\boldsymbol{N})(N+N+N)) = O(N^3)$
+
+After the first loop over the vertices in the main loop, the algorithm iterates over all the vertices to color all the vertices correctly in the main dictionary.
+The time complexity worst case at the end is $O((N \cdot (N+\boldsymbol{N})(N+N+N)) = O(N^3)$.
 

### 2016-04-06 paper/paper.tex
changes to graph and paper

@@
 \documentclass[twoside]{article}
+
+\usepackage{graphicx}
+
+\usepackage{listings}
+\lstset{
+breaklines=true,
+tabsize=2
+}
 
@@
 This paper will compare the time complexity of two color refinement algorithms and confirm them with computational results.
-The first algorithm uses the color refinement algorithm as shown by J.
-de Jong\cite{presentation:slidesPartI}, while the other algorithm is based on DFA minimization as shown by P.
+The first algorithm, henceforth referred to as the basic algorithm, uses the color refinement algorithm as shown by J.
+de Jong\cite{presentation:slidesPartI}, while the other algorithm, henceforth referred to as the fast algorithm, is based on DFA minimization as shown by P.
 Bonsma\cite{presentation:slidesPartIII}.
@@
 The aim of this research paper is to calculate and compare the worst-case time complexity of both color refinement algorithms and verifying them using computational results.
-These results will be used to examine the differences in computation time between the algorithms for different types of graphs.
+These results will be used to examine the differences in the growth of the computation time for a specific type of graph.
 
@@
 %% WIP %%
-The time complexity for both algorithms was calculated by hand.
+The time complexity of both algorithms was calculated by hand.
 To verify the results, a computational experiment was performed.
-In this experiment, both algorithms were tested with various graphs from the test set.
+In this experiment, both algorithms were tested with various sizes of a graph consisting of three paths.
+The paths are connected to each other on both of their ends.
+In the experiment the only thing changing is the length of these paths.
 
-The testing was all done on the same machine, under the same circumstances. Every test was performed at least three times in order to prevent outliers caused by external events.
+The testing was all done on the same machine, under the same circumstances. Every test was performed three times in order to prevent outliers caused by external events.
 
@@
 \label{methods:testing}
-In the experiment, both algorithms were tested with several sizes of the same type of graph, for various types of graphs. 
-% TODO Insert method of testing here
-Only the algorithms themselves were timed, initialization and other parts of the code were not included.
+In the experiment, both algorithms were tested with several sizes of the same type of graph, as described above.
+Each algorithm was timed three times.
+The lowest value of the three was selected as the most accurate, as the error margin for timing algorithms on a system is one-sided: it can only go slower than its optimal value.
+
+The python library used for timing the algorithm is \emph{timeit}. The manner in which is was used is the following:
+\begin{lstlisting}
+	min(Timer(algorithm_instance).repeat(3, 1))
+\end{lstlisting}
+
+Only the algorithms themselves and the initial coloring were timed, initialization and other parts of the code were not included.
 
@@
 % Table for results can be found at appendices
+The measured computation times are visualized in \autoref{fig:measurements}, the actual times can be found in \autoref{table:measurements}.
+
+\begin{figure}[H]
+	\includegraphics[width=\linewidth]{measurementsOnly}
+	\caption{The computation times of both the basic and fast color refinement algorithms for various sizes of the three paths graph.}
+	\label{fig:measurements}
+\end{figure}
 
@@
 %%
+
+\begin{figure}[H]
+	\includegraphics[width=\linewidth]{measurementsOnly}
+	\caption{The relative growth of measured computation time of the basic algorithm compared to the functions $f(x)=x^2$ and $f(x)=x^3$.}
+	\label{fig:basicGrowth}
+\end{figure}
+
+\begin{figure}[H]
+	\includegraphics[width=\linewidth]{measurementsOnly}
+	\caption{The relative growth of measured computation time of the fast algorithm compared to the functions $f(x)=x^2$ and $f(x)=n\cdot log(n)$.}
+	\label{fig:fastGrowth}
+\end{figure}
 
@@
 \begin{table}[h]\centering
-\label{app:tab:0}
 \begin{tabular}{|l|r@{.}l|r@{.}l|}
@@
 Graph	
-	& \multicolumn{2}{l|}{Algorithm A \emph{(s)}}
-	& \multicolumn{2}{l|}{Algorithm B \emph{(s)}} \\
+	& \multicolumn{2}{l|}{Basic \emph{(s)}}
+	& \multicolumn{2}{l|}{Fast \emph{(s)}} \\
 \hline
-graph\_{}type1\_{}16
-	& 0	& 001
-	& 0	& 005 \\
-graph\_{}type1\_{}64
-	& 0	& 1	
-	& 0	& 1 \\
-graph\_{}type1\_{}256
-	& 10	& 0	
-	& 2		& 0\\
-graph\_{}type1\_{}1024
-	& 1000	& 0 
-	& 40	& 0 \\
-graph\_{}type1\_{}4096
-	& 100000	& 0
-	& 800		& 0 \\
+threepaths\_{}5
+	& 0 & 00045
+	& 0	& 0 \\
+threepaths\_{}10
+	& 0 & 0017
+	& 0	& 0 \\
+threepaths\_{}20
+	& 0 & 0072
+	& 0	& 0 \\
+threepaths\_{}40
+	& 0 & 031
+	& 0	& 0 \\
+threepaths\_{}80
+	& 0 & 126
+	& 0	& 0 \\
+threepaths\_{}160
+	& 0 & 604
+	& 0	& 0 \\
+threepaths\_{}320
+	& 3 & 213
+	& 0	& 0 \\
+threepaths\_{}640
+	& 19 & 261
+	& 0	& 0 \\
+threepaths\_{}1280
+	& 127 & 567
+	& 0	& 0 \\
+threepaths\_{}2560
+	& 907 & 407
+	& 0	& 0 \\
+threepaths\_{}5120
+	& 5736 & 550
+	& 0	& 0 \\
+threepaths\_{}10240
+	& \multicolumn{2}{c|}{-}
+	& 0	& 0 \\
+
 \hline
 \end{tabular}
-\caption{This is a table}
+\caption{The measured computation times of the basic and fast color refinement algorithms for various sizes of the three paths graph.}
+\label{table:measurements}
 \end{table}

### 2016-04-06 graphs.m

@@ -27,16 +27,19 @@ timef2 = [ 0.04 0.09 0.13 0.3 1 3 12 ];
 p = semilogy(sizeb1, timeb1);
 p.Color = 'blue';
 p.Marker = '.';
-p.MarkerSize = 8;
+p.MarkerSize = 10;
+p.LineWidth = 1;
 
 hold on
-title('Computation Time of Basic Color Refinement for Threepaths')
-xlabel('Size \it (# vertices per path)')
-ylabel('Time \it (s)')
-xlabels = [ 5 320 640 1280 2560 5120 10240 ];
+title('Computation Time of Basic Color Refinement for Threepaths');
+xlabel('Size \it (# vertices per path)');
+ylabel('Time \it (s)');
+xlim([5 5120]);
+xlabels = [ 5 640 1280 2560 5120 10240 ];
 set(gca, 'Xtick', xlabels);
 set(gca, 'XtickLabel', xlabels);
 
+%%
 % Basic Slides w/out Adjacency
 p = semilogy(sizeb2, timeb2);
 p.Color = 'magenta';

### 2016-04-06 experiment/fastrefine.py
Shitty slow fast color

@@ -1,0 +1,188 @@
+from experiment.structures.basicgraphs import Graph
+from experiment.structures.doublylinkedlist import DoublyLinkedList
+
+
+def initial_coloring_fast(g: Graph):
+    g.queue = DoublyLinkedList()
+    g.color_classes = {}  # (key, value) = (color_num, ColorClass Object)
+
+    g.available_colors = DoublyLinkedList()  # using pop left, we can efficiently grab the lowest unused color.
+    for i in range(len(g.vertices())):
+        g.available_colors.append(i)
+
+    del_colors = []  # local list to check if color has been removed already.
+    queue_pointer = {}  # (color, dllnode)
+    biggest_cc = (0, 0)  # (color, size)
+    for v in g.vertices():
+        # initial coloring
+        v.colornum = v.degree()
+
+        # update available colors list and update color class
+        if v.colornum in del_colors:  # color has been seen before
+            # add vertex to color class
+            cc = g.color_classes[v.colornum]
+            dll_node = cc.vertices.append(v)
+            if len(cc.vertices) > biggest_cc[1]:  # calling len() on a dll, is O(1).
+                biggest_cc = (v.colornum, len(cc.vertices))
+        else:  # color hasn't been seen before.
+            # update available colors
+            g.available_colors.remove(v.colornum)
+            del_colors.append(v.colornum)
+
+            # create color_dict entry
+            cc = ColorClass(in_queue=True, color_num=v.colornum)
+            dll_node = cc.vertices.append(v)
+            g.color_classes[v.colornum] = cc
+
+            # add color to queue
+            queue_node = g.queue.append(v.colornum)
+            queue_pointer[v.colornum] = queue_node
+        v.dll_node = dll_node
+    # remove biggest color class from queue.
+    queue_pointer[biggest_cc[0]].delete_self()
+    g.color_classes[biggest_cc[0]].in_queue = False
+
+
+def refine_fast(g: Graph):
+    if hasattr(g, 'queue') and hasattr(g, 'color_classes'):
+        if hasattr(g, 'available_colors'):
+            while len(g.queue) > 0:
+                #print(g.queue)
+                c_color = g.queue.pop_left()
+
+                for (c_prime, cc_prime) in g.color_classes.copy().items():
+                    if c_prime == c_color or len(cc_prime.vertices) == 1:
+                        """
+                        a class doesn't refine itself.
+                        and a class of 1 node can't be refined.
+                        """
+                        continue
+                    else:
+                        """
[…]

### 2016-04-06 experiment/graphutil.py

@@ -1,0 +1,16 @@
+from experiment.structures.basicgraphs import Graph
+
+
+def prerequisites(g: Graph, h: Graph) -> bool:
+    return len(g.vertices()) == len(h.vertices()) and len(g.edges()) == len(h.edges())
+
+
+def disjoint_union(g: Graph, h: Graph, unsafe: bool = False) -> Graph:
+    # Stolen from blackboard.
+    vertex_map = {}
+    r = Graph(unsafe=unsafe)
+    for v in g.vertices() + h.vertices():
+        vertex_map[v] = r.add_vertex()
+    for e in g.edges() + h.edges():
+        r.add_edge(vertex_map[e.tail()], vertex_map[e.head()])
+    return r

### 2016-04-06 experiment/main.py

@@ -1,19 +1,26 @@
 import sys
+import time
 from enum import Enum
 
-from experiment.algorithm import *
+from experiment.fastrefine import *
+from experiment.graphutil import prerequisites, disjoint_union
 from experiment.io import graphIO as io
+from experiment.slowrefine import *
 
 # graph_file = '../graphs/colorref_smallexample_6_15.grl'
+# graph_file = '../graphs/colorref_smallexample_2_49.grl'
+
+
 # graph_file = '../graphs/colorref_largeexample_4_1026.grl'
 # graph_file = '../graphs/torus144.grl'
 # graph_file = '../graphs/trees90.grl'
-graph_file = '../graphs/products72.grl'
+# graph_file = '../graphs/products72.grl'
 
 
 # graph_file = '../graphs/cographs1.grl' # TODO: goes out of range, colors aren't freed or something?
-# raph_file = '../graphs/bigtrees1.grl'
+# graph_file = '../graphs/bigtrees1.grl'
 # graph_file = '../graphs/wheelstar12.grl'
+graph_file = '../graphs/threepaths640.gr'
 
 
 def iso_check(g: Graph, h: Graph):
@@ -37,7 +44,7 @@ class Result(Enum):
     ISOMORHP, NOT_ISOMORPH, DIFFERENCE_IN_EDGES_OR_VERTICES, UNDETERMINED = range(4)
 
 
-def main():
+def slow():
     graphs = io.loadgraph(filename=graph_file, readlist=True)[0]
     isomorphisms = []
     number_of_graphs = len(graphs)
@@ -67,10 +74,52 @@ def main():
             print(i)
 
 
-def refine_only():
-    graph = graphs = io.loadgraph(filename=graph_file, readlist=True)[0][0]
+def fast():
+    graphs = io.loadgraph(filename=graph_file, readlist=True)[0]
+    """
+    for i in range(0, len(graphs)):
+        for j in range(i + 1, len(graphs)):
+    """
+    # g = disjoint_union(graphs[0], graphs[1], unsafe=True)
+    g = graphs[0]
+    h = copy.copy(g)
+    f = copy.copy(g)
+
+    h1 = time.time()
+    initial_coloring(h)
+    h2 = time.time()
+    color_refine(h)
+    h3 = time.time()
+
+    hi = h2 - h1
+    hr = h3 - h2
+    print("Slow: " + str(hi) + " | " + str(hr))
[…]

### 2016-04-06 experiment/slowrefine.py

@@ -133,18 +133,3 @@ def isomorph_test(g: Graph):
             return branching(g)
     else:
         return False
-
-
-def prerequisites(g: Graph, h: Graph) -> bool:
-    return len(g.vertices()) == len(h.vertices()) and len(g.edges()) == len(h.edges())
-
-
-def disjoint_union(g: Graph, h: Graph, unsafe: bool = False) -> Graph:
-    # Stolen from blackboard.
-    vertex_map = {}
-    r = Graph(unsafe=unsafe)
-    for v in g.vertices() + h.vertices():
-        vertex_map[v] = r.add_vertex()
-    for e in g.edges() + h.edges():
-        r.add_edge(vertex_map[e.tail()], vertex_map[e.head()])
-    return r

### 2016-04-06 experiment/structures/doublylinkedlist.py

@@ -1,21 +1,55 @@
-class _DoublyNode:
-    def __init__(self, data, prev, next_):
+class DoublyNode:
+    def __init__(self, data, dll, prev=None, next_=None):
+        self._dll = dll
         self.data = data
         self.prev = prev
         self.next = next_
         pass
 
+    def delete_self(self):
+        is_head = self == self._dll._head
+        is_tail = self == self._dll._tail
+
+
+        if is_head and is_tail:
+            self._dll._head = None
+            self._dll._tail = None
+        elif is_head:
+            self.next.prev = None
+            self._dll._head = self.next
+            self.next = None
+        elif is_tail:
+            self.prev.next = None
+            self._dll._tail = self.prev
+            self.prev = None
+        else:
+            self.prev.next = self.next
+            self.next.prev = self.prev
+
+        self._dll._len -= 1  # ugly but efficient.
+        pass
+
+    def get_dll(self):
+        return self._dll
+
     pass
 
 
 class DoublyLinkedList:
-    def __init__(self):
+    def __init__(self, data : [] = None):
         self._head = None
         self._tail = None
+        self._len = 0
+        self._current = self._head
+
+        if data is not None:
+            for i in data:
+                self.append(i)
         pass
 
     def append(self, data):
-        new_node = _DoublyNode(data, None, None)
+        new_node = DoublyNode(data, self)
+        self._len += 1
         if self._head is None:
             # empty list
             self._head = self._tail = new_node
@@ -25,13 +59,24 @@ class DoublyLinkedList:
             new_node.prev = self._tail
             # update tail
             self._tail = new_node
-        pass
+        return new_node
+
+    def push(self, data):
+        return self.append(data)
+
+    def push_left(self, data):
+        new_node = DoublyNode(data, dll=self, next_=self._head)
+        self._head.prev = new_node
+        self._head = new_node
+        self._len += 1
+        return new_node
 
     def remove(self, data, all_instances: bool = False):
         pointer = self._head
         p = None
         while pointer is not None:
             if pointer.data == data:
+                self._len -= 1
                 # first element is head
                 if pointer.prev is None:
                     self._head = pointer.next
@@ -54,6 +99,29 @@ class DoublyLinkedList:
                 pointer = pointer.next
         pass
 
+    def pop(self):
+        node = self._tail
+        self._tail = node.prev
+        node.prev = None
+        self._tail.next = None
+        self._len -= 1
+        return node.data
+
+    def pop_left(self):
[…]

### 2016-04-06 experiment/structures/doublylinkedlist.py
dll

@@ -47,6 +47,19 @@ class DoublyLinkedList:
                 self.append(i)
         pass
 
+    def __iter__(self):
+        self._current = self._head
+        return self
+
+    def __next__(self):
+        if self._current is None:
+            self._current = self._head
+            raise StopIteration
+        else:
+            r = self._current
+            self._current = self._current.next
+            return r.data
+
     def append(self, data):
         new_node = DoublyNode(data, self)
         self._len += 1

### 2016-04-06 colorRefinement.py
stuff

@@ -12,16 +12,16 @@ from makegraphs import disjointunion
 
 def fastrefine(g, colordictarg=-1, startcolor=-1, preproc=False):
     if colordictarg is -1:
-        if g.colordict == -1:
-            if not preproc:
+        # if g.colordict == -1:
+            # if not preproc:
                 colordict = degcolordict(g)
-            else:
-                colordict = preproccolordict(g)
-        else:
-            if not preproc:
-                colordict = g.colordict
-            else:
-                colordict = preproccolordict
+            # else:
+            #     colordict = preproccolordict(g)
+        # else:
+        #     if not preproc:
+        #         colordict = g.colordict
+        #     else:
+        #         colordict = preproccolordict
     else:
         colordict = colordictarg
 
@@ -431,25 +431,26 @@ def componentpreproc(graphlisturl):
             isomorphisms.append(grl.index(graph))
     print('shit finished, isomorphisms:', isomorphisms)
 
-    elapsed_time = time.clock() - start_time
-    print('Components: {0:.4f} sec'.format(elapsed_time))
+    # elapsed_time = time.clock() - start_time
+    # print('Components: {0:.4f} sec'.format(elapsed_time))
 
 
-start_time = time.clock()
+# start_time = time.clock()
 
-sys.setrecursionlimit(10000)
+# sys.setrecursionlimit(10000)
 
-comparepreproc("GI_TestInstancesWeek1/torus24.grl", False)  # GI for large 4 1026
-comparepreproc("GI_TestInstancesWeek1/trees90.grl", False)  # #aut for torus72
-comparepreproc("GI_TestInstancesWeek1/products72.grl", False)  # #aut for product72
-comparepreproc("GI_TestInstancesWeek1/cographs1.grl", False)  # GI for cubes6
-compare("GI_TestInstancesWeek1/bigtrees1.grl", True)  # GI for bigtrees3
-#
+# compare("Basic/basicGI1.grl", True)
+# compare("Basic/basicGI2.grl", True)
+# compare("Basic/basicGI3.grl", True)
+# compare("Basic/basicGIAut.grl", False)
+# compare("Basic/basicAut1.gr", False)
+# compare("Basic/basicAut2.gr", False)
+
 # comparepreproc("GI_TestInstancesWeek1/cographs1.grl")  # GI for cographs1 with preprocessing
[…]

### 2016-04-06 tests.py

@@ -1,0 +1,97 @@
+from timeit import *
+
+from colorRefinement import *
+from graphIO import loadgraph
+
+threepaths = []
+
+
+def main():
+    path = 'threepaths/threepaths'
+    threepaths.append(loadgraph(path + '5.gr'))
+    threepaths.append(loadgraph(path + '10.gr'))
+    threepaths.append(loadgraph(path + '20.gr'))
+    threepaths.append(loadgraph(path + '40.gr'))
+    threepaths.append(loadgraph(path + '80.gr'))
+    threepaths.append(loadgraph(path + '160.gr'))
+    threepaths.append(loadgraph(path + '320.gr'))
+    threepaths.append(loadgraph(path + '640.gr'))
+    threepaths.append(loadgraph(path + '1280.gr'))
+    threepaths.append(loadgraph(path + '2560.gr'))
+    threepaths.append(loadgraph(path + '5120.gr'))
+    threepaths.append(loadgraph(path + '10240.gr'))
+    print(min(Timer(t0).repeat(3, 1)))
+    print(min(Timer(t1).repeat(3, 1)))
+    print(min(Timer(t2).repeat(3, 1)))
+    print(min(Timer(t3).repeat(3, 1)))
+    print(min(Timer(t4).repeat(3, 1)))
+    print(min(Timer(t5).repeat(3, 1)))
+    print(min(Timer(t6).repeat(3, 1)))
+    print(min(Timer(t7).repeat(3, 1)))
+    print(min(Timer(t8).repeat(3, 1)))
+    print(min(Timer(t9).repeat(3, 1)))
+    print(min(Timer(t10).repeat(3, 1)))
+    print(min(Timer(t11).repeat(3, 1)))
+    # print(Timer(t0).repeat(3, 1))
+    # print(Timer(t1).repeat(3, 1))
+    # print(Timer(t2).repeat(3, 1))
+    # print(Timer(t3).repeat(3, 1))
+    # print(Timer(t4).repeat(3, 1))
+    # print(Timer(t5).repeat(3, 1))
+    # print(Timer(t6).repeat(3, 1))
+    # print(Timer(t7).repeat(3, 1))
+    # print(Timer(t8).repeat(3, 1))
+    # print(Timer(t9).repeat(3, 1))
+    # print(Timer(t10).repeat(3, 1))
+    # print(Timer(t11).repeat(3, 1))
+
+
+def t0():
[…]

### 2016-04-06 experiment/algorithms/slowrefine.py
Final experiment

@@ -1,5 +1,3 @@
-import copy
-
 from experiment.structures.basicgraphs import *
 
 
@@ -79,57 +77,3 @@ def color_refine(g: Graph):
         return
     else:
         raise GraphError("Graph has no color dictionary")
-
-
-def branching(g: Graph):
-    possible = []
-    g.colors_freq = {}
-    for v in g.vertices():
-        if v.colornum in g.colors_freq:
-            g.colors_freq[v.colornum].append(v)
-
-            if len(g.colors_freq[v.colornum]) >= 4:
-                if v.colornum not in possible:
-                    possible.append(v.colornum)
-        else:
-            g.colors_freq[v.colornum] = [v]
-
-    if possible:  # empty list is false
-        for c in possible:
-            cmpl = g.number_of_vertices() / 2
-            branches = []  # label numbers of possible branches
-            branch_label = None
-            for w in g.colors_freq[c]:
-                l = w.label()
-                if l < cmpl:
-                    branches.append(l)
-                else:
-                    branch_label = l
-
-            branch_color = new_color(g)
-            for b in branches:
-                h = copy.deepcopy(g)
-                bv = h[branch_label]
-                bv.colornum = branch_color
-                h[b].colornum = branch_color
-                h.color_dict[branch_color] = (bv.degree(), create_nbs_dict(bv))
-                color_refine(h)
-                return isomorph_test(h)
-
-
-def isomorph_test(g: Graph):
-    colors0 = [0] * g.number_of_vertices()
-    colors1 = [0] * g.number_of_vertices()
-    cmpl = g.number_of_vertices() / 2
-    for i in g.vertices():
-        if i.label() < cmpl:
-            colors0[i.colornum] += 1
-        else:
-            colors1[i.colornum] += 1
-    if colors0 == colors1:
-        if max(colors0) == 1:
-            return True
-        else:
-            return branching(g)
-    else:
-        return False

### 2016-04-06 experiment/fastrefine.py

@@ -1,188 +1,0 @@
-from experiment.structures.basicgraphs import Graph
-from experiment.structures.doublylinkedlist import DoublyLinkedList
-
-
-def initial_coloring_fast(g: Graph):
-    g.queue = DoublyLinkedList()
-    g.color_classes = {}  # (key, value) = (color_num, ColorClass Object)
-
-    g.available_colors = DoublyLinkedList()  # using pop left, we can efficiently grab the lowest unused color.
-    for i in range(len(g.vertices())):
-        g.available_colors.append(i)
-
-    del_colors = []  # local list to check if color has been removed already.
-    queue_pointer = {}  # (color, dllnode)
-    biggest_cc = (0, 0)  # (color, size)
-    for v in g.vertices():
-        # initial coloring
-        v.colornum = v.degree()
-
-        # update available colors list and update color class
-        if v.colornum in del_colors:  # color has been seen before
-            # add vertex to color class
-            cc = g.color_classes[v.colornum]
-            dll_node = cc.vertices.append(v)
-            if len(cc.vertices) > biggest_cc[1]:  # calling len() on a dll, is O(1).
-                biggest_cc = (v.colornum, len(cc.vertices))
-        else:  # color hasn't been seen before.
-            # update available colors
-            g.available_colors.remove(v.colornum)
-            del_colors.append(v.colornum)
-
-            # create color_dict entry
-            cc = ColorClass(in_queue=True, color_num=v.colornum)
-            dll_node = cc.vertices.append(v)
-            g.color_classes[v.colornum] = cc
-
-            # add color to queue
-            queue_node = g.queue.append(v.colornum)
-            queue_pointer[v.colornum] = queue_node
-        v.dll_node = dll_node
-    # remove biggest color class from queue.
-    queue_pointer[biggest_cc[0]].delete_self()
-    g.color_classes[biggest_cc[0]].in_queue = False
-
-
-def refine_fast(g: Graph):
-    if hasattr(g, 'queue') and hasattr(g, 'color_classes'):
-        if hasattr(g, 'available_colors'):
-            while len(g.queue) > 0:
-                #print(g.queue)
-                c_color = g.queue.pop_left()
-
-                for (c_prime, cc_prime) in g.color_classes.copy().items():
-                    if c_prime == c_color or len(cc_prime.vertices) == 1:
-                        """
-                        a class doesn't refine itself.
-                        and a class of 1 node can't be refined.
-                        """
-                        continue
-                    else:
-                        """
[…]

### 2016-04-06 experiment/graphutil.py

@@ -1,16 +1,0 @@
-from experiment.structures.basicgraphs import Graph
-
-
-def prerequisites(g: Graph, h: Graph) -> bool:
-    return len(g.vertices()) == len(h.vertices()) and len(g.edges()) == len(h.edges())
-
-
-def disjoint_union(g: Graph, h: Graph, unsafe: bool = False) -> Graph:
-    # Stolen from blackboard.
-    vertex_map = {}
-    r = Graph(unsafe=unsafe)
-    for v in g.vertices() + h.vertices():
-        vertex_map[v] = r.add_vertex()
-    for e in g.edges() + h.edges():
-        r.add_edge(vertex_map[e.tail()], vertex_map[e.head()])
-    return r

### 2016-04-06 experiment/main.py

@@ -1,125 +1,59 @@
-import sys
+import os
 import time
-from enum import Enum
 
-from experiment.fastrefine import *
-from experiment.graphutil import prerequisites, disjoint_union
+from experiment.algorithms.fastrefine import *
+from experiment.algorithms.slowrefine import *
 from experiment.io import graphIO as io
-from experiment.slowrefine import *
 
-# graph_file = '../graphs/colorref_smallexample_6_15.grl'
-# graph_file = '../graphs/colorref_smallexample_2_49.grl'
+tp5 = 'testinstances/threepaths5.gr'
+tp80 = 'testinstances/threepaths80.gr'
+tp320 = 'testinstances/threepaths320.gr'
+tp640 = 'testinstances/threepaths640.gr'
+tp1280 = 'testinstances/threepaths1280.gr'
+tp5120 = 'testinstances/threepaths5120.gr'
+tp10240 = 'testinstances/threepaths10240.gr'
 
-
-# graph_file = '../graphs/colorref_largeexample_4_1026.grl'
-# graph_file = '../graphs/torus144.grl'
-# graph_file = '../graphs/trees90.grl'
-# graph_file = '../graphs/products72.grl'
-
-
-# graph_file = '../graphs/cographs1.grl' # TODO: goes out of range, colors aren't freed or something?
-# graph_file = '../graphs/bigtrees1.grl'
-# graph_file = '../graphs/wheelstar12.grl'
-graph_file = '../graphs/threepaths640.gr'
+graph_file_list = [tp5, tp80, tp320, tp640, tp1280, tp5120, tp10240]
+SLOW = False  # Should we run color refinement using the slow algorithm
 
 
-def iso_check(g: Graph, h: Graph):
-    if prerequisites(g, h):
-        i = disjoint_union(g, h, unsafe=True)
-        initial_coloring(i)
-        color_refine(i)
-        r = isomorph_test(i)
+def experiment():
+    for graph_file in graph_file_list:
+        print("Graph: " + str(os.path.basename(graph_file)))
+        print("Type | Time initial | Time refinement | Color classes")
+        graphs = io.loadgraph(filename=graph_file, readlist=True)[0]
+        g = graphs[0]
 
-        if r is None:
-            return Result.UNDETERMINED
-        elif r:
-            return Result.ISOMORHP
-        else:
-            return Result.NOT_ISOMORPH
-    else:
[…]

### 2016-04-06 experiment/structures/doublylinkedlist.py

@@ -10,7 +10,6 @@ class DoublyNode:
         is_head = self == self._dll._head
         is_tail = self == self._dll._tail
 
-
         if is_head and is_tail:
             self._dll._head = None
             self._dll._tail = None
@@ -36,7 +35,7 @@ class DoublyNode:
 
 
 class DoublyLinkedList:
-    def __init__(self, data : [] = None):
+    def __init__(self, data: [] = None):
         self._head = None
         self._tail = None
         self._len = 0
@@ -146,4 +145,3 @@ class DoublyLinkedList:
             r += str(pointer.data)
             pointer = pointer.next
         return r + "]"
-

### 2016-04-07 graphs.m
graphs.m update

@@ -14,12 +14,18 @@ sizeb3 = [ 160 320 640 1280 ];
 timeb3 = [ 0.3 2 17 125 ];
 
 % Fast Color Refinement
-
+% Measurement 1
+sizef1 = [ 5 10 20 40 80 160 320 640 1280 2560 5120 10240 20480];
+timef1 = [ 0.00058746337890625 0.0008630752563476562 0.0017960071563720703 0.0044100284576416016 0.007059335708618164 0.014141559600830078 0.028933286666870117 0.058183908462524414 0.11742138862609863 0.26059627532958984 0.491408348083496 0.994426965713501 2.088636636734009 ];
 % Slides w/ adjacency
 sizef2 = [ 160 320 640 1280 2560 5120 10240 ];
 timef2 = [ 0.04 0.09 0.13 0.3 1 3 12 ];
+% Measurement 2
+sizef3 = size;
+timef3 = [ 0.000124264999612933 0.000342142000590683 0.00100838100024702 0.00321382000038283 0.0112542269998812 0.0411697640001875 0.157870802999241 0.60982414099999 2.40300816400031 9.60912720700071 42.0685733139998 188.312010114 ];
 
-%% Computation Time Subplot
+
+%% Basic Computation Time Subplot
 
 %subplot(2,1,1);
 
@@ -39,7 +45,54 @@ xlabels = [ 5 640 1280 2560 5120 10240 ];
 set(gca, 'Xtick', xlabels);
 set(gca, 'XtickLabel', xlabels);
 
-%%
+% x^2
+x = 0:5:10240;
+p = semilogy(x, x.^2);
+p.Color = 'black';
+
+% x^3
+p = semilogy(x, x.^3);
+p.Color = 'red';
+
+legend('color refinement', 'n^2', 'n^3');
+legend('Location','southeast');
+legend('boxon');
+
+%% Fast Computation Time Subplot
+
+%subplot(2,1,1);
+
+% Basic Measurement 1
+p = semilogy(sizef1, timef1);
+p.Color = 'blue';
+p.Marker = '.';
+p.MarkerSize = 10;
+p.LineWidth = 1;
+
+hold on
+title('Computation Time of Fast Color Refinement for Threepaths');
+xlabel('Size \it (# vertices per path)');
+ylabel('Time \it (s)');
+xlim([5 20480]);
+xlabels = [ 5 2560 5120 10240  20480];
[…]

### 2016-04-07 paper/paper.tex
updated graphs and added correct figures

@@
 \label{methods}
-%% WIP %%
 The time complexity of both algorithms was calculated by hand.
@@
 \label{algs}
-%% TODO %%
+%% WIP %%
 % Describe algorithms, show concise calculations of time complexity
@@
 % Table for results can be found at appendices
-The measured computation times are visualized in \autoref{fig:measurements}, the actual times can be found in \autoref{table:measurements}.
+The measured computation times are visualized in \autoref{fig:basiconly} and \autoref{fig:fastonly}, the actual times can be found in \autoref{table:measurements}.
 
 \begin{figure}[H]
-	\includegraphics[width=\linewidth]{measurementsOnly}
-	\caption{The computation times of both the basic and fast color refinement algorithms for various sizes of the three paths graph.}
-	\label{fig:measurements}
+	\includegraphics[width=\linewidth]{basiconly}
+	\caption{The computation times of the basic color refinement algorithm for various sizes of the three paths graph.}
+	\label{fig:basiconly}
+\end{figure}
+
+\begin{figure}[H]
+	\includegraphics[width=\linewidth]{fastonly}
+	\caption{The computation times of the fast color refinement algorithm for various sizes of the three paths graph.}
+	\label{fig:fastonly}
 \end{figure}
@@
 \begin{figure}[H]
-	\includegraphics[width=\linewidth]{measurementsOnly}
-	\caption{The relative growth of measured computation time of the basic algorithm compared to the functions $f(x)=x^2$ and $f(x)=x^3$.}
-	\label{fig:basicGrowth}
+	\includegraphics[width=\linewidth]{basicfunctions}
+	\caption{The computation times of the basic color refinement algorithm and the functions $f(n)=n^2$ and $f(n)=n^3$.}
+	\label{fig:basicfunctions}
 \end{figure}
@@
 \begin{figure}[H]
-	\includegraphics[width=\linewidth]{measurementsOnly}
-	\caption{The relative growth of measured computation time of the fast algorithm compared to the functions $f(x)=x^2$ and $f(x)=n\cdot log(n)$.}
-	\label{fig:fastGrowth}
+	\includegraphics[width=\linewidth]{basicdiff}
+	\caption{The relative growth of measured computation time of the basic algorithm compared to the functions $f(n)=n^2$ and $f(n)=n^3$.}
+	\label{fig:basicgrowth}
+\end{figure}
+
+\begin{figure}[H]
+	\includegraphics[width=\linewidth]{fastfunctions}
+	\caption{The computation times of the fast color refinement algorithm and the functions $f(x)=n\cdot log(n)$ and $f(x)=n$.}
+	\label{fig:fastgrowth}
+\end{figure}
+
+\begin{figure}[H]
+	\includegraphics[width=\linewidth]{fastdiff}
+	\caption{The relative growth of measured computation time of the fast algorithm compared to the functions $f(x)=n\cdot log(n)$ and $f(x)=n$.}
+	\label{fig:fastfunctions}
 \end{figure}
@@
 	& 0 & 00045
-	& 0	& 0 \\
+	& 0	& 00059 \\
 threepaths\_{}10
 	& 0 & 0017
-	& 0	& 0 \\
+	& 0	& 00086 \\
 threepaths\_{}20
 	& 0 & 0072
-	& 0	& 0 \\
+	& 0	& 0018 \\
 threepaths\_{}40
 	& 0 & 031
-	& 0	& 0 \\
+	& 0	& 0044 \\
 threepaths\_{}80
 	& 0 & 126
-	& 0	& 0 \\
+	& 0	& 0071 \\
 threepaths\_{}160
 	& 0 & 604
-	& 0	& 0 \\
+	& 0	& 014 \\
 threepaths\_{}320
 	& 3 & 213
-	& 0	& 0 \\
+	& 0	& 029 \\
 threepaths\_{}640
 	& 19 & 261
-	& 0	& 0 \\
+	& 0	& 058 \\
 threepaths\_{}1280
 	& 127 & 567
-	& 0	& 0 \\
+	& 0	& 117 \\
 threepaths\_{}2560
 	& 907 & 407
-	& 0	& 0 \\
+	& 0	& 261 \\
 threepaths\_{}5120
 	& 5736 & 550
-	& 0	& 0 \\
+	& 0	& 491 \\
 threepaths\_{}10240
 	& \multicolumn{2}{c|}{-}
-	& 0	& 0 \\
+	& 0	& 994 \\
+threepaths\_{}10240
+	& \multicolumn{2}{c|}{-}
+	& 2	& 089 \\
 

### 2016-04-07 graphs.m

@@ -78,9 +78,9 @@ xlabels = [ 5 2560 5120 10240  20480];
 set(gca, 'Xtick', xlabels);
 set(gca, 'XtickLabel', xlabels);
 
-% log2n
+% n
 x = 0:5:20480;
-p = semilogy(x, log2(x));
+p = semilogy(x, x);
 p.Color = 'black';
 
 % nlog2n
@@ -88,7 +88,7 @@ x = 0:5:20480;
 p = semilogy(x, x.*log2(x));
 p.Color = 'red';
 
-legend('color refinement', 'log_2n', 'n\cdotlog_2n');
+legend('color refinement', 'n', 'n\cdotlog_2n');
 legend('Location','southeast');
 legend('boxon');
 
@@ -127,18 +127,19 @@ legend('Measurements', 'Slides w/out adjacency', 'Slides w/ adjacency', 'Fast sl
 
 %% Relative Growth Basic Subplot
 m2 = sizeb1.^2;
-diff2 = m2./timeb1;
+diff2 = m2-timeb1;
 m3 = sizeb1.^3;
 diff3 = m3./timeb1;
-diff = [ diff3.' diff2.' ];
+%diff = [ diff3.' diff2.' ];
+diff = diff2.';
 
 %subplot(2,1,2);
 figure
 
 bar_handle = bar(diff, 'grouped');
-set(gca, 'YScale', 'log');
+set(gca, 'YScale', 'linear');
 set(bar_handle(1), 'FaceColor', 'red')
-set(bar_handle(2), 'FaceColor', 'black')
+%set(bar_handle(2), 'FaceColor', 'black')
 
 set(gca, 'XtickLabel', sizeb1);
 
@@ -160,7 +161,7 @@ diff6 = m6./timef1;
 m7 = sizef1.^2.*log2(sizef1);
 diff7 = m7./timef1;
 m8 = log(sizef1);
-diff8 = m8./timef1;
+diff8 = sizef1./timef1;
 diff = [ diff4.' diff8.'];
 
 %subplot(2,1,2);
@@ -184,6 +185,6 @@ title('Relative Differences between Fast Algorithm and Functions')
 xlabel('Size \it (# vertices per path)')
 ylabel('Function divided by computation time')
 
-legend('n\cdotlog_2n', 'log_2n');
+legend('n\cdotlog_2n', 'n');
 legend('Location','northwest');
 legend('boxon');

### 2016-04-07 paper/paper.tex
sectioning in section 5 and 6
some sort of formatting, the horizontal scroll bar annoyed me
Paragraph 4 structure overhaul

@@
-%----------------------------------------------------------------------------------------
+%------------------------------------------------
 %	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
-%----------------------------------------------------------------------------------------
+%------------------------------------------------
 
@@
 
-%----------------------------------------------------------------------------------------
+%------------------------------------------------
 %	TITLE SECTION
-%----------------------------------------------------------------------------------------
+%------------------------------------------------
 
@@
 
-%----------------------------------------------------------------------------------------
+%------------------------------------------------
 
@@
 
-%----------------------------------------------------------------------------------------
+%------------------------------------------------
 %	ABSTRACT
-%----------------------------------------------------------------------------------------
+%------------------------------------------------
 
@@
 
-%----------------------------------------------------------------------------------------
+%------------------------------------------------
 %	ARTICLE CONTENTS
-%----------------------------------------------------------------------------------------
+%------------------------------------------------
 
@@
 
-The python library used for timing the algorithm is \emph{timeit}. The manner in which is was used is the following:
+The python library used for timing the algorithm is \emph{timeit}. The manner in which it was used is the following:
 \begin{lstlisting}
@@
 %%
-For the paper, two different color refinement methods were implemented.
-The basic one gives all vertices in a graph with the same color but differently colored neighborhoods a separate color.
-This is done iteratively until the graph is stably colored.
-A coloring $\alpha$ of $G$ is stable if for all $u, v \in V(G)$ it holds that: if $\alpha(u) = \alpha(v)$, then $u$ and $v$ have identically colored neighborhoods.
+In the following subsections, we give a general overview of the algorithms and analyze them:
+The computation time of a algorithms increases for iterations over the vertices in a graph.
+To calculate the worst-case time complexity of the algorithms, the number and places of the iterations over the vertices are analyzed.
+
+The input and output for the algorithms are the same.
+The input is a graph with a possibly unstable  coloring.
+The output is graph with stable coloring.
+However, the graphs do have a different structure to support their respective algorithms refinement operation.
+
+A coloring $\alpha$ of a graph $G$ is called stable if for all $u, v \in V(G)$ it holds that: if $\alpha(u) = \alpha(v)$, then $u$ and $v$ have identically colored neighborhoods.
+
+\subsection{Basic}
+The basic algorithm gives all vertices in a graph with the same color but differently colored neighborhoods a separate color.
+This is done iteratively until the graph has a stable coloring.
 The problem with this method is that all vertices are checked each iteration.
 
-With fast color refinement the colors that lead to refinements are tracked.(TODO: make better explanation) These colors are maintained in a queue.
-If a color class is being refined, the class is split into two different classes, with one class keeping the old color.
-The algorithm terminates with a stable coloring if no color class needs refinement, in other words the queue is empty.
-
-The initial coloring is done in a separate algorithm before the color refinement algorithms take place and is therefore the same for both the fast and basic implementation.
-
-The computation time of our algorithms increases for iterations over the vertices in a graph.
-To calculate the time complexity of the algorithms, the number and places of the iterations over the vertices are analyzed.
-
 The basic implementation has several places where iterations over vertices take place.
-The code contains one main loop that has a worst case of N iterations.
+The code contains one main loop that has a worst-case of N iterations.
 Thereby the time complexity so far is $O(\boldsymbol{N})$, where $N = V(G)$.
@@
 If the the configuration is not yet know, a new color is assigned.
-In worst case this new color operation costs N.
+In worst-case this new color operation costs N.
 The time complexity is $O((N \cdot N)(N+N+\boldsymbol{N}))$ so far.
@@
 After the first loop over the vertices in the main loop, the algorithm iterates over all the vertices to color all the vertices correctly in the main dictionary.
-The time complexity worst case at the end is $O((N \cdot (N+\boldsymbol{N})(N+N+N)) = O(N^3)$.
+The time complexity worst-case at the end is $O((N \cdot (N+\boldsymbol{N})(N+N+N)) = O(N^3)$.
+
+\subsection{Fast}
+With fast color refinement the colors that lead to refinements are tracked.(TODO: make better explanation) These colors are maintained in a queue.
+If a color class is being refined, the class is split into two different classes, with one class keeping the old color.
+The algorithm terminates with a stable coloring if no color class needs refinement, in other words the queue is empty.
+%% TODO %%
 
@@
 
+\subsection{Basic Algorithm}
+\autoref{fig:basiconly} shows the measure computation times of the basic color refinement algorithm for the three paths graphs of sizes 5 through 5120. The larger sizes of this graph
+
 \begin{figure}[H]
@@
 \end{figure}
+
+\subsection{Fast Algorithm}
 
@@
 
+\subsection{Basic Algorithm}
+
 \begin{figure}[H]
@@
 
+\subsection{Comparison}
+
 %------------------------------------------------
@@
 
-%----------------------------------------------------------------------------------------
+%------------------------------------------------
 %	REFERENCE LIST
-%----------------------------------------------------------------------------------------
+%------------------------------------------------
 \label{references}
@@
 
-%----------------------------------------------------------------------------------------
+%------------------------------------------------
 %	APPENDICES
-%----------------------------------------------------------------------------------------
+%------------------------------------------------
 %% TODO %%
@@
 \end{appendices}
-%----------------------------------------------------------------------------------------
+%------------------------------------------------
 

### 2016-04-07 paper/paper.tex
temp commit
added some more description to results, probably needs more, but my mind is kind of blank atm

@@
 \label{results}
-%% TODO %%
+%% WIP %%
 % Give and describe results
@@
 \subsection{Basic Algorithm}
-\autoref{fig:basiconly} shows the measure computation times of the basic color refinement algorithm for the three paths graphs of sizes 5 through 5120. The larger sizes of this graph
+\autoref{fig:basiconly} shows the measured computation times of the basic color refinement algorithm for the three paths graphs of sizes 5 through 5120.
+The larger sizes of this graph would have taken too long to compute within reasonable time.
 
@@
 \subsection{Fast Algorithm}
+\autoref{fig:fastonly} shows the measured computation times of the fast color refinement algorithm for the three paths graphs of sizes 5 through 20480.
+Larger variants of the graph were not available and therefore not used.
 
@@
 \label{disc}
-%% TODO %%
+%% WIP %%
 % Describe the effectiveness of the algorithms on various types of graphs, and examine the differences

### 2016-04-07 paper/paper.tex
Small changes 4

--- 2016-04-07 paper/paper.tex
@@
-%------------------------------------------------
+%----------------------------------------------------------------------------------------
 %	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
-%------------------------------------------------
+%----------------------------------------------------------------------------------------
 
@@
 
-%------------------------------------------------
+%----------------------------------------------------------------------------------------
 %	TITLE SECTION
-%------------------------------------------------
+%----------------------------------------------------------------------------------------
 
@@
 
-%------------------------------------------------
+%----------------------------------------------------------------------------------------
 
@@
 
-%------------------------------------------------
+%----------------------------------------------------------------------------------------
 %	ABSTRACT
-%------------------------------------------------
+%----------------------------------------------------------------------------------------
 
@@
 
-%------------------------------------------------
+%----------------------------------------------------------------------------------------
 %	ARTICLE CONTENTS
-%------------------------------------------------
+%----------------------------------------------------------------------------------------
 
@@
 \label{methods}
+%% WIP %%
 The time complexity of both algorithms was calculated by hand.
@@
 \label{algs}
-%% WIP %%
+%% TODO %%
 % Describe algorithms, show concise calculations of time complexity
@@
 The input is a graph with a possibly unstable  coloring.
-The output is graph with stable coloring.
-However, the graphs do have a different structure to support their respective algorithms refinement operation.
+The output is a graph with stable coloring.
+However, the graphs do have a different structure to support their respective algorithm's refinement operation.
+We do not discuss the initialisation of these structures.
 
@@
 This is done iteratively until the graph has a stable coloring.
-The problem with this method is that all vertices are checked each iteration.
+The determining factor of this method is that all vertices are checked each iteration.
 
@@
 % Table for results can be found at appendices
-The measured computation times are visualized in \autoref{fig:basiconly} and \autoref{fig:fastonly}, the actual times can be found in \autoref{table:measurements}.
-
-\subsection{Basic Algorithm}
-\autoref{fig:basiconly} shows the measure computation times of the basic color refinement algorithm for the three paths graphs of sizes 5 through 5120. The larger sizes of this graph
+The measured computation times are visualized in \autoref{fig:measurements}, the actual times can be found in \autoref{table:measurements}.
 
 \begin{figure}[H]
-	\includegraphics[width=\linewidth]{basiconly}
-	\caption{The computation times of the basic color refinement algorithm for various sizes of the three paths graph.}
-	\label{fig:basiconly}
-\end{figure}
-
-\subsection{Fast Algorithm}
-
-\begin{figure}[H]
-	\includegraphics[width=\linewidth]{fastonly}
-	\caption{The computation times of the fast color refinement algorithm for various sizes of the three paths graph.}
-	\label{fig:fastonly}
+	\includegraphics[width=\linewidth]{measurementsOnly}
+	\caption{The computation times of both the basic and fast color refinement algorithms for various sizes of the three paths graph.}
+	\label{fig:measurements}
 \end{figure}
@@
 
-\subsection{Basic Algorithm}
-
 \begin{figure}[H]
-	\includegraphics[width=\linewidth]{basicfunctions}
-	\caption{The computation times of the basic color refinement algorithm and the functions $f(n)=n^2$ and $f(n)=n^3$.}
-	\label{fig:basicfunctions}
+	\includegraphics[width=\linewidth]{measurementsOnly}
+	\caption{The relative growth of measured computation time of the basic algorithm compared to the functions $f(x)=x^2$ and $f(x)=x^3$.}
+	\label{fig:basicGrowth}
 \end{figure}
@@
 \begin{figure}[H]
-	\includegraphics[width=\linewidth]{basicdiff}
-	\caption{The relative growth of measured computation time of the basic algorithm compared to the functions $f(n)=n^2$ and $f(n)=n^3$.}
-	\label{fig:basicgrowth}
+	\includegraphics[width=\linewidth]{measurementsOnly}
+	\caption{The relative growth of measured computation time of the fast algorithm compared to the functions $f(x)=x^2$ and $f(x)=n\cdot log(n)$.}
+	\label{fig:fastGrowth}
 \end{figure}
-
-\subsection{Fast Algorithm}
-
-\begin{figure}[H]
-	\includegraphics[width=\linewidth]{fastfunctions}
-	\caption{The computation times of the fast color refinement algorithm and the functions $f(x)=n\cdot log(n)$ and $f(x)=n$.}
-	\label{fig:fastgrowth}
-\end{figure}
-
-\begin{figure}[H]
-	\includegraphics[width=\linewidth]{fastdiff}
-	\caption{The relative growth of measured computation time of the fast algorithm compared to the functions $f(x)=n\cdot log(n)$ and $f(x)=n$.}
-	\label{fig:fastfunctions}
-\end{figure}
-
-\subsection{Comparison}
 
@@
 
-%------------------------------------------------
+%----------------------------------------------------------------------------------------
 %	REFERENCE LIST
-%------------------------------------------------
+%----------------------------------------------------------------------------------------
 \label{references}
@@
 
-%------------------------------------------------
+%----------------------------------------------------------------------------------------
 %	APPENDICES
-%------------------------------------------------
+%----------------------------------------------------------------------------------------
 %% TODO %%
@@
 	& 0 & 00045
-	& 0	& 00059 \\
+	& 0	& 0 \\
 threepaths\_{}10
 	& 0 & 0017
-	& 0	& 00086 \\
+	& 0	& 0 \\
 threepaths\_{}20
 	& 0 & 0072
-	& 0	& 0018 \\
+	& 0	& 0 \\
 threepaths\_{}40
 	& 0 & 031
-	& 0	& 0044 \\
+	& 0	& 0 \\
 threepaths\_{}80
 	& 0 & 126
-	& 0	& 0071 \\
+	& 0	& 0 \\
 threepaths\_{}160
 	& 0 & 604
-	& 0	& 014 \\
+	& 0	& 0 \\
 threepaths\_{}320
 	& 3 & 213
-	& 0	& 029 \\
+	& 0	& 0 \\
 threepaths\_{}640
 	& 19 & 261
-	& 0	& 058 \\
+	& 0	& 0 \\
 threepaths\_{}1280
 	& 127 & 567
-	& 0	& 117 \\
+	& 0	& 0 \\
 threepaths\_{}2560
 	& 907 & 407
-	& 0	& 261 \\
+	& 0	& 0 \\
 threepaths\_{}5120
 	& 5736 & 550
-	& 0	& 491 \\
+	& 0	& 0 \\
 threepaths\_{}10240
 	& \multicolumn{2}{c|}{-}
-	& 0	& 994 \\
-threepaths\_{}10240
-	& \multicolumn{2}{c|}{-}
-	& 2	& 089 \\
+	& 0	& 0 \\
 
@@
 \end{appendices}
-%------------------------------------------------
+%----------------------------------------------------------------------------------------
 

### 2016-04-07 paper/paper.tex

@@
 % Table for results can be found at appendices
-The measured computation times are visualized in \autoref{fig:measurements}, the actual times can be found in \autoref{table:measurements}.
+The measured computation times are visualized in \autoref{fig:basiconly} and \autoref{fig:fastonly}, the actual times can be found in \autoref{table:measurements}.
+
+\subsection{Basic Algorithm}
+\autoref{fig:basiconly} shows the measured computation times of the basic color refinement algorithm for the three paths graphs of sizes 5 through 5120.
+The larger sizes of this graph would have taken too long to compute within reasonable time.
+
+\begin{figure}[H]
+	\includegraphics[width=\linewidth]{basiconly}
+	\caption{The computation times of the basic color refinement algorithm for various sizes of the three paths graph.}
+	\label{fig:basiconly}
+\end{figure}
+
+\subsection{Fast Algorithm}
+\autoref{fig:fastonly} shows the measured computation times of the fast color refinement algorithm for the three paths graphs of sizes 5 through 20480.
+Larger variants of the graph were not available and therefore not used.
 

### 2016-04-07 paper/paper.tex
idk anymore
fingers crossed

--- 2016-04-07 paper/paper.tex
@@
 %%
+To verify the results
 

### 2016-04-07 paper/paper.tex
discussion

@@
 %%
-To verify the results
+To verify the results of the time complexity calculations, the measurements have to be compared to functions of the expected time complexity.
+To more clearly see the relation between the two, diagrams with the relative size of the two were created.
 
 \subsection{Basic Algorithm}
+As the calculated worst-case time complexity of the basic color refinement algorithm is $O(N^3)$, \autoref{fig:basicfunctions} shows the measurements and the function $f(n)=x^3$.
+\autoref{fig:basicgrowth} then plots the relative size of the two of on a logarithmic scale.
+It shows that $f(n)=x^3$ grows significantly faster than the measurements.
+Looking for a better fit, $f(n)=n^2$ was added to both diagrams.
+\autoref{fig:fastgrowth} indicates
 
@@
 	\caption{The computation times of the fast color refinement algorithm and the functions $f(x)=n\cdot log(n)$ and $f(x)=n$.}
-	\label{fig:fastgrowth}
+	\label{fig:fastfunctions}
 \end{figure}
@@
 	\caption{The relative growth of measured computation time of the fast algorithm compared to the functions $f(x)=n\cdot log(n)$ and $f(x)=n$.}
-	\label{fig:fastfunctions}
+	\label{fig:fastgrowth}
 \end{figure}

### 2016-04-07 paper/paper.tex
discussion

@@
 As the calculated worst-case time complexity of the basic color refinement algorithm is $O(N^3)$, \autoref{fig:basicfunctions} shows the measurements and the function $f(n)=x^3$.
-\autoref{fig:basicgrowth} then plots the relative size of the two of on a logarithmic scale.
+\autoref{fig:basicgrowth} then plots the relative size of the two on a logarithmic scale.
 It shows that $f(n)=x^3$ grows significantly faster than the measurements.
 Looking for a better fit, $f(n)=n^2$ was added to both diagrams.
-\autoref{fig:fastgrowth} indicates
+\autoref{fig:basicgrowth} indicates that the relative size is slowly decreasing and the growth of the measurements is approaching the growth of $f(n)=n^2$.
 
@@
 \subsection{Fast Algorithm}
+As the calculated worst-case time complexity of the fast color refinement algorithm is $O(M\cdot{}log(N)$, \autoref{fig:fastfunctions} shows the measurements and the function $f(n)=n\cdot{}log(n)$.
+\autoref{fig:fastgrowth} then plots the relative size of the two on a logarithmic scale.
+It shows that $f(n)=n\cdot{}log(n)$ grows faster than the measurements.
+Looking for a better fit, $f(n)=n$ was added to both diagrams.
+\autoref{fig:fastgrowth} indicates that $f(n)=n$ and the measurements are growing at the same rate, which implies that the time complexity for this specific type of graph might be closer to $\Theta(N)$.
 

### 2016-04-07 paper/paper.tex
minor fix in discussion
abstract

@@
 
-\noindent This will be the abstract of the paper
+\noindent The worst-case asymptotic time complexity for a basic and fast color refinement algorithm are calculated to be $O(N^3)$ and $O(M\cdot{}log(N)$ respectively.
+Computational results for a three paths type graph indicate different time complexities, at least for that specific type of graph.
+It can be concluded that either the algorithms are faster for the three paths graphs, or the calculations for the time complexity are incorrect.
 
@@
 
-\subsection{Comparison}
-
 %------------------------------------------------

### 2016-04-07 paper/paper.tex
Conclusion

@@
 \label{concl}
+The most obvious conclusion we draw from the worst-case complexity and computational results is that fast refinement is a faster algorithm than basic refinement.
+However, when we compare the computational results with their calculated worst-case complexity we observe some differences.
+
+The complexity of basic refinement we determined is $O(N^3)$. When we look at the relative growth, it seems that the algorithm approaches $O(N^2)$. There could be two causes:
+\begin{compactitem}
+	\item The determined complexity of $O(N^3)$ was faulty.
+	\item The type of graphs used, threegraphs, aproach a $O(N^2)$ complexity because of their structure. $O(N^3)$ was determined as worst-case.
+\end{compactitem}
+
+The complexity of fast refinement we determined is $O(MlogN)$.
+When we look at the relative growth, it seems that the algorithm approaches a linear complexity.
+We attribute this to the structure of the type of graphs used, threegraphs.
+$O(MlogN)$ was determined as worst-case.
 %% TODO %%

### 2016-04-07 paper/paper.tex
references and stuff
stuk van huub + small fixes

@@
 
+The fast color refinement algorithm used for this paper is made by W. van Veelen, N. Goldsmid, E. Khan, and E. Koster \cite{misc:group4}, because the original implementation had some issues.
+
 %------------------------------------------------
@@
 \subsection{Fast}
-With fast color refinement the colors that lead to refinements are tracked.(TODO: make better explanation) These colors are maintained in a queue.
+With fast color refinement the colors that lead to refinements are tracked.
+These colors are maintained in a queue.
 If a color class is being refined, the class is split into two different classes, with one class keeping the old color.
 The algorithm terminates with a stable coloring if no color class needs refinement, in other words the queue is empty.
-%% TODO %%
+
+Fast color refinement is also analyzed.
+$N$ is again defined as $N=|V(G)|$.
+Determined is the number of times a vertex $v$ is refined in its worst-case.
+The color class of $v$ splits at least in half if it is getting refined.
+The split class with the most vertices in it keeps the original color.
+If vertex $v$ stays in this original class, $v$ is not refined by definition.
+It follows that if $v$ gets refined, the degree of the previous color class is at most two times the degree of its new class.
+Let $k$ be the times $v$ refines in the algorithm, then $2^k$ is the degree of $v$'s original color class.
+The degree of the color class is worst-case $N$ if all vertices have the same amount of edges, so worst-case $k=log_2 N$.
+As shown by P.
+Bonsma in one of his lectures.
+\cite{presentation:slidesPartIII}
+
+With the number of edges in a graph times the number of vertices in that graph, the time complexity can be defined.\cite{presentation:slidesPartIII} The time complexity of fast color refinement is $O(m\cdot log n)$.
 

### 2016-04-08 paper/paper.tex
package
stuff

@@
 \usepackage{graphicx}
+
+\usepackage{algorithmicx}
 
@@
 \label{scope}
-The aim of this research paper is to calculate and compare the worst-case time complexity of both color refinement algorithms and verifying them using computational results.
+The aim of this research paper is to calculate and compare the worst-case asymptotic time complexity of both color refinement algorithms and verifying them using computational results.
 These results will be used to examine the differences in the growth of the computation time for a specific type of graph.
@@
 %%
-In the following subsections, we give a general overview of the algorithms and analyze them:
-The computation time of a algorithms increases for iterations over the vertices in a graph.
+In the following subsection, a general overview and an analysis of the algorithms is given:
+The computation time of the algorithms increases for iterations over the vertices in a graph.
 To calculate the worst-case time complexity of the algorithms, the number and places of the iterations over the vertices are analyzed.
@@
 The time complexity worst-case at the end is $O((N \cdot (N+\boldsymbol{N})(N+N+N)) = O(N^3)$.
+At the end, the worst-case time complexity is $O((N \cdot (N+\boldsymbol{N})(N+N+N)) = O(N^3)$.
 
 \subsection{Fast}
-With fast color refinement the colors that lead to refinements are tracked.
-These colors are maintained in a queue.
+With fast color refinement the colorsclasses that could be used to refinement other color classes are tracked by putting them in a queue.
 If a color class is being refined, the class is split into two different classes, with one class keeping the old color.
@@
 It follows that if $v$ gets refined, the degree of the previous color class is at most two times the degree of its new class.
-Let $k$ be the times $v$ refines in the algorithm, then $2^k$ is the degree of $v$'s original color class.
+Let $k$ be the times $v$ is refined in the algorithm, then $2^k$ is the degree of $v$'s original color class.
 The degree of the color class is worst-case $N$ if all vertices have the same amount of edges, so worst-case $k=log_2 N$.
@@
 \subsection{Basic Algorithm}
-\autoref{fig:basiconly} shows the measured computation times of the basic color refinement algorithm for the three paths graphs of sizes 5 through 5120.
+\autoref{fig:basiconly} shows the measured computation times of the basic color refinement algorithm for the three paths graphs with path lengths 5 through 5120.
 The larger sizes of this graph would have taken too long to compute within reasonable time.
 
-\begin{figure}[H]
-	\includegraphics[width=\linewidth]{basiconly}
-	\caption{The computation times of the basic color refinement algorithm for various sizes of the three paths graph.}
-	\label{fig:basiconly}
-\end{figure}
-
 \subsection{Fast Algorithm}
-\autoref{fig:fastonly} shows the measured computation times of the fast color refinement algorithm for the three paths graphs of sizes 5 through 20480.
+\autoref{fig:fastonly} shows the measured computation times of the fast color refinement algorithm for the three paths graphs with path lengths 5 through 20480.
 Larger variants of the graph were not available and therefore not used.
-
-\begin{figure}[H]
-	\includegraphics[width=\linewidth]{fastonly}
-	\caption{The computation times of the fast color refinement algorithm for various sizes of the three paths graph.}
-	\label{fig:fastonly}
-\end{figure}
 
@@
 \subsection{Basic Algorithm}
-As the calculated worst-case time complexity of the basic color refinement algorithm is $O(N^3)$, \autoref{fig:basicfunctions} shows the measurements and the function $f(n)=x^3$.
+As the calculated worst-case time complexity of the basic color refinement algorithm is $O(N^3)$, \autoref{fig:basicfunctions} shows the measurements and the function $f(n)=n^3$.
 \autoref{fig:basicgrowth} then plots the relative size of the two on a logarithmic scale.
-It shows that $f(n)=x^3$ grows significantly faster than the measurements.
+It shows that $f(n)=n^3$ grows significantly faster than the measurements.
 Looking for a better fit, $f(n)=n^2$ was added to both diagrams.
@@
 
-\begin{figure}[H]
-	\includegraphics[width=\linewidth]{basicfunctions}
-	\caption{The computation times of the basic color refinement algorithm and the functions $f(n)=n^2$ and $f(n)=n^3$.}
-	\label{fig:basicfunctions}
-\end{figure}
-
-\begin{figure}[H]
-	\includegraphics[width=\linewidth]{basicdiff}
-	\caption{The relative growth of measured computation time of the basic algorithm compared to the functions $f(n)=n^2$ and $f(n)=n^3$.}
-	\label{fig:basicgrowth}
-\end{figure}
-
 \subsection{Fast Algorithm}
-As the calculated worst-case time complexity of the fast color refinement algorithm is $O(M\cdot{}log(N)$, \autoref{fig:fastfunctions} shows the measurements and the function $f(n)=n\cdot{}log(n)$.
+As the calculated worst-case time complexity of the fast color refinement algorithm is $O(M\cdot{}logN)$, \autoref{fig:fastfunctions} shows the measurements and the function $f(n)=n\cdot{}log(n)$.
 \autoref{fig:fastgrowth} then plots the relative size of the two on a logarithmic scale.
@@
 
-\begin{figure}[H]
-	\includegraphics[width=\linewidth]{fastfunctions}
-	\caption{The computation times of the fast color refinement algorithm and the functions $f(x)=n\cdot log(n)$ and $f(x)=n$.}
-	\label{fig:fastfunctions}
-\end{figure}
-
-\begin{figure}[H]
-	\includegraphics[width=\linewidth]{fastdiff}
-	\caption{The relative growth of measured computation time of the fast algorithm compared to the functions $f(x)=n\cdot log(n)$ and $f(x)=n$.}
-	\label{fig:fastgrowth}
-\end{figure}
-
 %------------------------------------------------
@@
 \label{concl}
-The most obvious conclusion we draw from the worst-case complexity and computational results is that fast refinement is a faster algorithm than basic refinement.
-However, when we compare the computational results with their calculated worst-case complexity we observe some differences.
+The most obvious conclusion that can be drawn from the worst-case time complexity and computational results is that fast refinement is a faster algorithm than basic refinement.
+However, the computational results are compared with their calculated worst-case complexity some differences were observed.
 
-The complexity of basic refinement we determined is $O(N^3)$. When we look at the relative growth, it seems that the algorithm approaches $O(N^2)$. There could be two causes:
+The determined time complexity of basic refinement is $O(N^3)$. When looking at the relative growth, it seems that the algorithm approaches $O(N^2)$. There could be two causes:
 \begin{compactitem}
 	\item The determined complexity of $O(N^3)$ was faulty.
-	\item The type of graphs used, threegraphs, aproach a $O(N^2)$ complexity because of their structure. $O(N^3)$ was determined as worst-case.
+	\item The structure of the three paths graphs causes the basic algorithm to approach a $O(N^2)$ time complexity. $O(N^3)$ was determined as worst-case.
 \end{compactitem}
 
-The complexity of fast refinement we determined is $O(MlogN)$.
-When we look at the relative growth, it seems that the algorithm approaches a linear complexity.
-We attribute this to the structure of the type of graphs used, threegraphs.
+The determined time complexity of the fast algorithm is $O(MlogN)$.
+When looking at the relative growth, it seems that the algorithm approaches a linear complexity.
+This can be attributed to the structure of the type of graphs used, three paths.
 $O(MlogN)$ was determined as worst-case.
-%% TODO %%
-% Draw conlusions from both the results and discussion
-%%
 
@@
 %------------------------------------------------
+\section{Algorithms}
+\label{app:algs}
+
+%------------------------------------------------
+\section{Figures}
+\label{app:figs}
+\begin{multicols}{2}
+\begin{figure}[H]
+	\includegraphics[width=\linewidth]{basiconly}
+	\caption{The computation times of the basic color refinement algorithm for various sizes of the three paths graph.}
+	\label{fig:basiconly}
+\end{figure}
+
+\begin{figure}[H]
+	\includegraphics[width=\linewidth]{fastonly}
+	\caption{The computation times of the fast color refinement algorithm for various sizes of the three paths graph.}
+	\label{fig:fastonly}
+\end{figure}
+
+\begin{figure}[H]
+	\includegraphics[width=\linewidth]{basicfunctions}
+	\caption{The computation times of the basic color refinement algorithm and the functions $f(n)=n^2$ and $f(n)=n^3$.}
+	\label{fig:basicfunctions}
+\end{figure}
+
+\begin{figure}[H]
+	\includegraphics[width=\linewidth]{basicdiff}
+	\caption{The relative growth of measured computation time of the basic algorithm compared to the functions $f(n)=n^2$ and $f(n)=n^3$.}
+	\label{fig:basicgrowth}
+\end{figure}
+
+\begin{figure}[H]
+	\includegraphics[width=\linewidth]{fastfunctions}
+	\caption{The computation times of the fast color refinement algorithm and the functions $f(x)=n\cdot log(n)$ and $f(x)=n$.}
+	\label{fig:fastfunctions}
+\end{figure}
+
+\begin{figure}[H]
+	\includegraphics[width=\linewidth]{fastdiff}
+	\caption{The relative growth of measured computation time of the fast algorithm compared to the functions $f(x)=n\cdot log(n)$ and $f(x)=n$.}
+	\label{fig:fastgrowth}
+\end{figure}
+\end{multicols}
+
+%------------------------------------------------
+\newpage
 \section{Tables} %Change section name
-\label{app:tab}
-%% TODO %%
-% Fill table with actual values
-%%
-%% NOTE %%
-% All current values are fictional values for testing purposes only
-%%
+\label{app:tabs}
 \begin{table}[h]\centering

### 2016-04-08 paper/paper.tex
alg

@@
 
-\usepackage{algorithmicx}
+\usepackage{algorithm}
+\usepackage{algpseudocode}
 
@@
 \fancyfoot{} % Blank out the default footer
-\fancyhead[C]{W.G. Puttenstein, R.L.H. Fontein, H.D. van Wieren, T. Kerkhoven: \shorttitle} % Custom header text
+\fancyhead[C]{R.L.H. Fontein, H.D. van Wieren, T. Kerkhoven: \shorttitle} % Custom header text
 \fancyfoot[RO,LE]{\thepage} % Custom footer text
@@
 \large
-\textsc{Gerwin Puttenstein, Rick Fontein, Huub van Wieren, and Tim Kerkhoven}\\[2mm] % Your name
+\textsc{Rick Fontein, Huub van Wieren, and Tim Kerkhoven}\\[2mm] % Your name
 \normalsize University of Twente \\ % Your institution
-\normalsize \href{mailto:w.g.puttenstein@student.utwente.nl}{w.g.puttenstein@student.utwente.nl}, 
-\href{mailto:r.l.h.fontein@student.utwente.nl}{r.l.h.fontein@student.utwente.nl},\\
-\normalsize \href{mailto:h.d.vanwieren@student.utwente.nl}{h.d.vanwieren@student.utwente.nl},
-\href{mailto:t.kerkhoven@student.utwente.nl}{t.kerkhoven@student.utwente.nl}% Your email addresses
+\normalsize \href{mailto:r.l.h.fontein@student.utwente.nl}{r.l.h.fontein@student.utwente.nl},
+\href{mailto:h.d.vanwieren@student.utwente.nl}{h.d.vanwieren@student.utwente.nl},\\
+\normalsize \href{mailto:t.kerkhoven@student.utwente.nl}{t.kerkhoven@student.utwente.nl}% Your email addresses
 }
@@
 \label{app:algs}
+\begin{algorithm}
+    \begin{algorithmic}[1] % The number tells where the line numbering should start
+        \Procedure{BasicColorRefinement}{$G$}
+            \While{not stable} \Comment{Worst case loop over every vertex}
+            	 \For{every vertex $v \in G(V)$}
+            	 	\State calculate neighborhood $nbs$ of $v$
+            	  	\If {$nbs$ of $v$ has not changed}
+    \State pass
+    				\Else
+    					\For{every item in the dictionary}
+    						\State checks if the configuration is not yet known.
+    					\EndFor
+    					\If{the configuration is not yet known}
+    						\State assign a new color \
+    					\EndIf
+    				\EndIf
+            	  \EndFor
+				\For{every vertex in G}
+					\State update neighborhood dictionary
+				\EndFor
+            \EndWhile\label{euclidendwhile}
+            \State \textbf{return} stable colored graph
+        \EndProcedure
+    \end{algorithmic}
+    \caption{Basic color refinement}
+    \label{euclid}
+\end{algorithm}
 
@@
 %------------------------------------------------
-\newpage
+% \newpage
 \section{Tables} %Change section name

### 2016-04-08 paper/paper.tex
bla
Algorithm 2
formatting
Algorithm 1

@@
 \usepackage{algorithm}
-\usepackage{algpseudocode}
+\usepackage[noend]{algpseudocode}
 
@@
     \begin{algorithmic}[1] % The number tells where the line numbering should start
-        \Procedure{BasicColorRefinement}{$G$}
+        \Procedure{BasicColorRefinement}{Graph $G(V,E)$, Coloring $C$}
             \While{not stable} \Comment{Worst case loop over every vertex}
-            	 \For{every vertex $v \in G(V)$}
-            	 	\State calculate neighborhood $nbs$ of $v$
-            	  	\If {$nbs$ of $v$ has not changed}
-    \State pass
+            	 \ForAll{$v \in V$}
+            	 	\State calculate $v.nbs$ 
+            	  	\If{$v.nbs \neq Q[v.class]$} \Comment{nbs is different from previous iteration}
+            	  		\If{$v.nbs \in C$} \Comment{New configuration already has a new color}
+            	  			\State Assign $v$ this color
             	  		\Else
-    					\For{every item in the dictionary}
-    						\State checks if the configuration is not yet known.
-    					\EndFor
-    					\If{the configuration is not yet known}
-    						\State assign a new color \
+            	  			\State $v.class = newColor()$
+            	  			\State $C$.add($v.class$)            	  		
             	  		\EndIf
@@
             	  \EndFor
-				\For{every vertex in G}
-					\State update neighborhood dictionary
+				\ForAll{$v \in V)$}
+					\State update $C[v]$
 				\EndFor
-            \EndWhile\label{euclidendwhile}
-            \State \textbf{return} stable colored graph
+            \EndWhile
+        \EndProcedure \Comment{Graph $G$ now has a stable coloring}
+    \end{algorithmic}
+    \caption{Basic color refinement}
+    \label{alg:basic}
+\end{algorithm}
+
+\begin{algorithm}
+    \begin{algorithmic}[1]
+    	\Procedure{FastColorRefinement}{$G(V,E)$}
+    		\While{$|Q| > 1$}
+    			\State $w$ = $Q$.next()
+    			\State for $i \geq 0 \in \mathbb{Z}$ let $D_i$ denote the set of $v \in V$ with $i$ neigbours in $w$ 
+    			\For{$c \in C \setminus w$}
+    				\State let $X_i = D_i \cap c \neq \emptyset$
+    				\For{all $X_i$}
+    					\If{$max(X) == X_i$}
+    						\State $c = X_i$
+    					\Else
+    						\State add $X_i$ as new class to $C$
+    					\EndIf    					
+    				\EndFor
+    			\If{$c \in Q$}
+    				\State $X = X \setminus c$
+    			\EndIf
+    			\State $Q$.add($X \setminus \max (X$))
+    			\EndFor
+
+    		\EndWhile 
         \EndProcedure
     \end{algorithmic}
-    \caption{Basic color refinement}
-    \label{euclid}
+    \caption{Fast color refinement}
 \end{algorithm}

### 2016-04-08 paper/paper.tex
ref

@@
 \subsection{Basic}
-The basic algorithm gives all vertices in a graph with the same color but differently colored neighborhoods a separate color.
+The basic algorithm (Algorithm \autoref{alg:basic}) gives all vertices in a graph with the same color but differently colored neighborhoods a separate color.
 This is done iteratively until the graph has a stable coloring.
@@
 \subsection{Fast}
-With fast color refinement the colorsclasses that could be used to refinement other color classes are tracked by putting them in a queue.
+With fast color refinement (Algorithm \autoref{alg:fast}) the colorsclasses that could be used to refinement other color classes are tracked by putting them in a queue.
 If a color class is being refined, the class is split into two different classes, with one class keeping the old color.
@@
     \begin{algorithmic}[1] % The number tells where the line numbering should start
-        \Procedure{BasicColorRefinement}{Graph $G(V,E)$, Coloring $C$}
+        \Procedure{BasicColorRefinement}{Graph $G(V,E)$}
             \While{not stable} \Comment{Worst case loop over every vertex}
             	 \ForAll{$v \in V$}
-            	 	\State calculate $v.nbs$ 
-            	  	\If{$v.nbs \neq Q[v.class]$} \Comment{nbs is different from previous iteration}
-            	  		\If{$v.nbs \in C$} \Comment{New configuration already has a new color}
-            	  			\State Assign $v$ this color
+            	 	\State calculate neighborhood $nbs$ of $v$
+            	  	\If {$nbs$ of $v$ has not changed}
+    \State pass
     				\Else
-            	  			\State $v.class = newColor()$
-            	  			\State $C$.add($v.class$)            	  		
+    					\ForAll{$u \in V$}
+    						\State checks if the configuration is not yet known.
+    					\EndFor
+    					\If{the configuration is not yet known}
+    						\State assign a new color
     					\EndIf
@@
             	  \EndFor
-				\ForAll{$v \in V)$}
-					\State update $C[v]$
+				\ForAll{$v \in G(V)$}
+					\State update neighborhood dictionary
 				\EndFor
@@
     \caption{Fast color refinement}
+    \label{alg:fast}
 \end{algorithm}
@@
 %------------------------------------------------
+\newpage
 \section{Figures}
@@
 %------------------------------------------------
-% \newpage
+\newpage
 \section{Tables} %Change section name

### 2016-04-08 paper/paper.tex
final version

--- 2016-04-08 paper/paper.tex
@@
 The time complexity is now $O(N \cdot \boldsymbol{N})$.
-Each vertex is checked to see if it needs an update.
-If so, a new color dictionary is made and a new loop over the dictionary takes place.
-This makes the time complexity is $O(N \cdot N \cdot \boldsymbol{N})$ for making the dictionary and $O((N \cdot N)(N+\boldsymbol{N}))$ for iterating over the dictionary.
-Every color class in the dictionary is checked to see if the configuration has already been assigned a new color.
+The neighborhood of the vertex is calculated.
+If this neighborhood differs from the previous iteration, a new color dictionary is made is made in the source and a new loop over the dictionary takes place.
+This makes the time complexity $O(N \cdot N \cdot \boldsymbol{N})$ for making the dictionary and $O((N \cdot N)(N+\boldsymbol{N}))$ for iterating over the dictionary.
+Then every color class in the dictionary is checked to see if the configuration has already been assigned a new color.
 If the the configuration is not yet know, a new color is assigned.
-In worst-case this new color operation costs N.
+In worst-case the \emph{newColor()} function costs N.
 The time complexity is $O((N \cdot N)(N+N+\boldsymbol{N}))$ so far.
 
-After the first loop over the vertices in the main loop, the algorithm iterates over all the vertices to color all the vertices correctly in the main dictionary.
-The time complexity worst-case at the end is $O((N \cdot (N+\boldsymbol{N})(N+N+N)) = O(N^3)$.
-At the end, the worst-case time complexity is $O((N \cdot (N+\boldsymbol{N})(N+N+N)) = O(N^3)$.
+After the first loop over the vertices in the main loop, the algorithm iterates over all the vertices to update the coloring for all vertices.
+The worst-case time complexity at the end is $O((N \cdot (N+\boldsymbol{N})(N+N+N)) = O(N^3)$.
 
 \subsection{Fast}
-With fast color refinement the colorsclasses that could be used to refinement other color classes are tracked by putting them in a queue.
+With fast color refinement (Algorithm \autoref{alg:fast}) the color classes that could be used to refinement other color classes are tracked by putting them in a queue.
 If a color class is being refined, the class is split into two different classes, with one class keeping the old color.
